<!-- MS_Tab_Fig_Apndx-030-Table-1.qmd -->
<!-- Rich Jones --> 
<!-- 2025-07-22 -->

```{r}
# Study sample: all HRS/HCAP 2016 participants (R/MS_MAIN/MS_001_analysis_cohort.R).
study_sample <- analysis2_hcap %>% 
  dplyr::mutate(
    rHISP = dplyr::recode(HISPANIC,
                                      `0` = 0,
                                      `1` = 1,
                                      `2` = 1,
                                      `3` = 1,
                                      `5` = 5,
                                      .default = NA_real_),
    hcap_classification = factor(
      vs1hcapdxeap,
      levels = c(1, 2, 3),
      labels = c("Normal", "Mild cognitive impairment", "Dementia")
    ),
    langa_weir_classification = factor(
      cogfunction2016,
      levels = c(1, 2, 3),
      labels = c("Normal", "Cognitively impaired, not demented", "Dementia")
    )
  )

# GENDER
# HISPANIC
# RACE 
# SCHLYRS 
# PA019 (age)

```

<!-- Generate table inputs -->
```{r}
# library(forcats)
# library(rlang)
# library(dplyr)
# library(tibble)
summarize_variable <- function(data, var) {
  if (is.numeric(data[[var]])) {
    mean_val <- mean(data[[var]], na.rm = TRUE)
    sd_val <- sd(data[[var]], na.rm = TRUE)
    tibble::tibble(
      variable = var,
      value = "Mean (SD)",
      n_pct_or_mean_sd = sprintf("%.1f (%.1f)", mean_val, sd_val)
    )
  } else {
    data |>
      mutate(tmp = forcats::fct_infreq(as.factor(.data[[var]]))) |>  # preserve labels
      count(tmp) |>
      mutate(
        variable = var,
        value = as.character(tmp),
        percent = round(100 * n / sum(n), 1),
        n_pct_or_mean_sd = paste0(format(n, big.mark = ","), " (", formatC(percent, format = "f", digits = 1), "%)")
      ) |>
      select(variable, value, n_pct_or_mean_sd)
  }
}

label_categorical_vars <- function(df) {
  df |>
    dplyr::mutate(
      female = factor(female, 
                      levels = c(0, 1), 
                      labels = c("Male", "Female")),
      rHISP = factor(rHISP,
                     levels = c(0,1,5),
                     labels = c("Not obtained",
                                "Hispanic",
                                "Not Hispanic")),
      HISPANIC = factor(HISPANIC, 
                        levels = c(0, 1, 2, 3, 5),  
                        labels = c("Not obtained", 
                                   "Hispanic, Mexican", 
                                   "Hispanic, Other",
                                   "Hispanic, type unknown", 
                                   "Not Hispanic")),
      RACE = factor(RACE, 
                    levels = c(0, 1, 2, 7),
                    labels = c("Not collected", "White", "Black", "All other groups"))
    )
}

study_sample <- label_categorical_vars(study_sample)

vars <- c(
  "female", "rage", "RACE", "rHISP", "SCHLYRS",
  "hcap_classification", "langa_weir_classification", "jorm"
)

summary_final <- purrr::map_df(vars, summarize_variable, data = study_sample) %>%
  mutate(value = case_when(is.na(value) ~ "Blank, partial interview", 
                           !is.na(value) ~ value)) %>%
  rename(study_sample = n_pct_or_mean_sd) %>%
  arrange(factor(variable, levels = vars), value)

missing_summary <- tibble::tibble(
  variable = c("langa_weir_classification", "jorm"),
  value = "Missing",
  study_sample = format(
    c(
      sum(is.na(study_sample$langa_weir_classification)),
      sum(is.na(study_sample$jorm))
    ),
    big.mark = ","
  )
)

summary_final <- dplyr::bind_rows(summary_final, missing_summary)

# Create a lookup table for variable labels and types
var_labels <- tibble::tibble(
  variable = c(
    "female", "rage", "RACE", "rHISP", "SCHLYRS",
    "hcap_classification", "langa_weir_classification", "jorm"
  ),
  label = c(
    "Sex (N (%))",
    "Age in years (Mean (SD))",
    "Race (N (%))",
    "Hispanic origin (N (%))",
    "Years of education (Mean (SD))",
    "HRS/HCAP Classification (N (%))",
    "Langa-Weir Classification (N (%))",
    "Jorm IQCODE score from Core (Mean (SD))"
  ),
  type = c(
    "categorical", "continuous", "categorical", "categorical", "continuous",
    "categorical", "categorical", "continuous"
  )
)
```

```{r}
# Join labels to summary_final
summary_final <- summary_final %>%
  left_join(var_labels, by = "variable")

# Create a new label_combined column
summary_final <- summary_final %>%
  mutate(label_combined = dplyr::case_when(
    type == "continuous" & value == "Mean (SD)" ~ label,
    variable == "jorm" & value == "Missing" ~ "Jorm IQCODE score from Core (missing)",
    value == "Missing" ~ "   Missing",
    type == "categorical" ~ paste0("   ", value),
    TRUE ~ NA_character_
  ))

# Insert a header row before each categorical variable
insert_headers <- summary_final %>%
  filter(type == "categorical") %>%
  distinct(variable, label) %>%
  mutate(label_combined = label, study_sample = NA)

summary_final <- bind_rows(summary_final, insert_headers) 

summary_final <- summary_final %>%
  mutate(
    sortorder = sub(" .*", "", sub("^\\s+", "", study_sample)),
    sortorder = as.numeric(gsub(",", "", sortorder)),
    sortorder = ifelse(is.na(sortorder), 999999, sortorder)
  )

summary_final <- summary_final %>%
  mutate(variable = factor(variable, levels = vars)) %>%
  arrange(variable, desc(sortorder)) %>% 
  select(label_combined, study_sample) 

```


```{r}
#| echo: false 
#| message: false
#| warning: false 

flextable::flextable(summary_final) |>
  flextable::set_header_labels(
    label_combined = "Characteristic",
    study_sample = paste0(
      "HRS/HCAP 2016 participants\n(N = ",
      format(nrow(study_sample), big.mark = ","), ")"
    )
  ) |>
  flextable::add_header_lines("Table 1. Participant Characteristics") |>
  flextable::add_footer_lines(
    "Source: R/MS_MAIN/MS_001_analysis_cohort.R; R_objects/A7_005_hrs16_merged.rds."
  ) |>
  flextable::add_footer_lines(
    "Jorm IQCODE was collected in the HRS Core interview, not HCAP, and only from participants who did not complete Core cognitive testing."
  ) |>
  flextable::autofit()  |>
  flextable::fontsize(size = 10, part = "all") |>  # match body font size
  flextable::font(fontname = "Arial", part = "all")  
``` 


::: {custom-style="Footnote text"}
Notes: The study sample includes all `r format(nrow(study_sample), big.mark = ",")` HRS/HCAP 2016 participants, including participants without Core cognitive testing who were classified from informant (Jorm IQCODE) data. Race and ethnicity data were gathered via self-report at the time of first interview and are considered to be markers of exposure to evolving systems of racism, not as a proxy for genetic variation or any other biological variable. Race was self-selected by participants at the time of the first interview from a list of options defined by the 2000 US Census criteria. Other race includes a pooled group of participants who identified as American Indian or Alaska Native, Asian, Native Hawaiian or Pacific Islander, or another self-described race, consolidated due to small sample sizes and risk of identification.
:::
