
```{r}
hrs16_cog <- readRDS(fs::path(r_objects_folder, "011_hrs16_cog.rds"))

```

# Recode demographic variables

This section shows the recoding of the demographic variables.

Age came from the variable `PA019` and no recoding was done.

Years of education came from the variable `SCHLYRS` and no recoding was done.  Five participants had their values changed based on RNJ's code.

Female came from the variable `GENDER`.

```{r}
#| tbl-cap: "Coding of female indicator"
#| label: tbl-female

hrs16_cog %>%
  mutate(GENDER = factor(GENDER, levels = c(1, 2),  labels = c("1.  Male", "2.  Female")) ) %>%
  QSPtools::checkvar(female, GENDER) %>%
  gt::gt()
```

Black came from the variables `RACE` and `HISPANIC`.

```{r}
#| tbl-cap: "Coding of Black or African-American indicator"
#| label: tbl-black

hrs16_cog %>%
  mutate(
    RACE = factor(RACE, levels = c(0, 1, 2, 7), 
                  labels = c("0.  Not obtained", "1.  White/Caucasian", 
                             "2.  Black or African American", "7.  Other")),
    HISPANIC = factor(HISPANIC, levels = c(0, 1, 2, 3, 5),  
                      labels = c("0.  Not obtained", "1.  Hispanic, Mexican", "2.  Hispanic, Other",
                                 "3.  Hispanic, type unknown", "5.  Non-Hispanic"))) %>%
  QSPtools::checkvar(black, RACE, HISPANIC) %>%
  gt::gt()
```

Hispanic came from the variable `HISPANIC`.

```{r}
#| tbl-cap: "Coding of Hispanic indicator"
#| label: tbl-hisp

hrs16_cog %>%
  mutate(
    HISPANIC = factor(HISPANIC, levels = c(0, 1, 2, 3, 5),  
                      labels = c("0.  Not obtained", "1.  Hispanic, Mexican", "2.  Hispanic, Other", 
                                 "3.  Hispanic, type unknown", "5.  Non-Hispanic"))) %>%
  QSPtools::checkvar(hisp, HISPANIC) %>%
  gt::gt()
```
