```{r}
hrs16_iadl <- readRDS(fs::path(r_objects_folder, "012_hrs16_iadl.rds"))

```

# Recode the functional impairment items

This section shows the recoding of the IADL items.

There are 11 IADL items that are considered to measure functional impairment.

-   Item responses of 1 (Yes) and 6 (Can't do) are recoded as "impairment"

-   Item responses of 5 (No), 7 (Don't do), 8 (Don't know) and 9 (Refused) are recoded as "no impairment"

Two of the items - "Difficulty taking medication" and "Do you think you would have difficulty taking medication" are logically dependent and are merged into one item

The sum of these 10 recoded IADL items is the functional impairment score

```{r}
#| tbl-cap: "ADL: Difficulty Dressing recoding"
#| label: tbl-dressing

hrs16_iadl %>%
  labelled::to_factor() %>%
  gtsummary::tbl_cross(PG014, rPG014) %>%
  my_gtsummary()
```

```{r}
#| tbl-cap: "ADL: Difficulty Bathing recoding"
#| label: tbl-bathing

hrs16_iadl %>%
  labelled::to_factor() %>%
  gtsummary::tbl_cross(PG021, rPG021) %>%
  my_gtsummary()
```

```{r}
#| tbl-cap: "ADL: Difficulty Eating recoding"
#| label: tbl-eating

hrs16_iadl %>%
  labelled::to_factor() %>%
  gtsummary::tbl_cross(PG023, rPG023) %>%
  my_gtsummary()
```

```{r}
#| tbl-cap: "ADL: Difficulty Using Toilet recoding"
#| label: tbl-toilet

hrs16_iadl %>%
  labelled::to_factor() %>%
  gtsummary::tbl_cross(PG030, rPG030) %>%
  my_gtsummary()
```

```{r}
#| tbl-cap: "ADL: Difficulty Using Maps recoding"
#| label: tbl-maps

hrs16_iadl %>%
  labelled::to_factor() %>%
  gtsummary::tbl_cross(PG040, rPG040) %>%
  my_gtsummary()
```

```{r}
#| tbl-cap: "IADL: Difficulty Meal Prep recoding"
#| label: tbl-meals

hrs16_iadl %>%
  labelled::to_factor() %>%
  gtsummary::tbl_cross(PG041, rPG041) %>%
  my_gtsummary()
```

```{r}
#| tbl-cap: "IADL: Difficulty Grocery Shopping recoding"
#| label: tbl-grocery

hrs16_iadl %>%
  labelled::to_factor() %>%
  gtsummary::tbl_cross(PG044, rPG044) %>%
  my_gtsummary()
```

```{r}
#| tbl-cap: "IADL: Difficulty Making Phone Calls recoding"
#| label: tbl-phone

hrs16_iadl %>%
  labelled::to_factor() %>%
  gtsummary::tbl_cross(PG047, rPG047) %>%
  my_gtsummary()
```

```{r}
#| tbl-cap: "IADL: Difficulty Taking Meds recoding"
#| label: tbl-meds

QSPtools::checkvar(hrs16_iadl, rPG050, PG050, PG051) %>%
  gt::gt() %>%
  my_gt()
```

```{r}
#| tbl-cap: "IADL: Difficulty Managing Money Calls recoding"
#| label: tbl-money

hrs16_iadl %>%
  labelled::to_factor() %>%
  gtsummary::tbl_cross(PG059, rPG059) %>%
  my_gtsummary()
```

```{r}
#| tbl-cap: "Descriptive statistics of recoded functional impairment items"
#| label: tbl-desc-iadls

hrs16_iadl %>%
  select(starts_with("r"), iadl_imp) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(iadl_imp) ~ "categorical"
  ) %>%
  my_gtsummary()


```
