

```{r}
hrs16_func <- readRDS(fs::path(r_objects_folder, "013_hrs16_func.rds"))
hrshcap <- readRDS(fs::path(r_objects_folder, "005_hrshcap.rds"))


```

# Recode the informant based items

This section shows the recoding of the Jorm items and the subjective cognitive impairment item.

There are 16 Jorm items.

-   For each Jorm item, three HRS items need to be combined to recreate the Jorm item

The average of these 16 Jorms items is the score

The subjective cognitive impairment item was recoded so that "no change" and "improvement" was combined into one category.



```{r}
#| tbl-cap: "Jorm: Remembering things about family"
#| label: tbl-jorm1

hrs16_func %>%
  QSPtools::checkvar(jorm1, PD506, PD507, PD508) %>%
  gt::gt() %>%
  my_gt()
```

```{r}
#| tbl-cap: "Jorm: Remembering things that happened recently"
#| label: tbl-jorm2

hrs16_func %>%
  QSPtools::checkvar(jorm2, PD509, PD510, PD511) %>%
  gt::gt() %>%
  my_gt()
```

```{r}
#| tbl-cap: "Jorm: Recalling conversations a few day later"
#| label: tbl-jorm3

hrs16_func %>%
  QSPtools::checkvar(jorm3, PD512, PD513, PD514) %>%
  gt::gt() %>%
  my_gt()
```

```{r}
#| tbl-cap: "Jorm: Remembering telephone number"
#| label: tbl-jorm4

hrs16_func %>%
  QSPtools::checkvar(jorm4, PD515, PD516, PD517) %>%
  gt::gt() %>%
  my_gt()
```

```{r}
#| tbl-cap: "Jorm: Remembering day and month"
#| label: tbl-jorm5

hrs16_func %>%
  QSPtools::checkvar(jorm5, PD518, PD519, PD520) %>%
  gt::gt() %>%
  my_gt()
```

```{r}
#| tbl-cap: "Jorm: Remembering where things are kept"
#| label: tbl-jorm6

hrs16_func %>%
  QSPtools::checkvar(jorm6, PD521, PD522, PD523) %>%
  gt::gt() %>%
  my_gt()
```

```{r}
#| tbl-cap: "Jorm: Remembering where to find things"
#| label: tbl-jorm7

hrs16_func %>%
  QSPtools::checkvar(jorm7, PD524, PD525, PD526) %>%
  gt::gt() %>%
  my_gt()
```

```{r}
#| tbl-cap: "Jorm: Knowing how to work familar machines around the house"
#| label: tbl-jorm8

hrs16_func %>%
  QSPtools::checkvar(jorm8, PD527, PD528, PD529) %>%
  gt::gt() %>%
  my_gt()
```

```{r}
#| tbl-cap: "Jorm: Learning to use a new gadget"
#| label: tbl-jorm9

hrs16_func %>%
  QSPtools::checkvar(jorm9, PD530, PD531, PD532) %>%
  gt::gt() %>%
  my_gt()
```

```{r}
#| tbl-cap: "Jorm: Learning new things in general"
#| label: tbl-jorm10

hrs16_func %>%
  QSPtools::checkvar(jorm10, PD533, PD534, PD535) %>%
  gt::gt() %>%
  my_gt()
```

```{r}
#| tbl-cap: "Jorm: Following a story in a book or on TV"
#| label: tbl-jorm11

hrs16_func %>%
  QSPtools::checkvar(jorm11, PD536, PD537, PD538) %>%
  gt::gt() %>%
  my_gt()
```

```{r}
#| tbl-cap: "Jorm: Making decisions on everyday matters"
#| label: tbl-jorm12

hrs16_func %>%
  QSPtools::checkvar(jorm12, PD539, PD540, PD541) %>%
  gt::gt() %>%
  my_gt()
```

```{r}
#| tbl-cap: "Jorm: Handling money for shopping"
#| label: tbl-jorm13

hrs16_func %>%
  QSPtools::checkvar(jorm13, PD542, PD543, PD544) %>%
  gt::gt() %>%
  my_gt()
```

```{r}
#| tbl-cap: "Jorm: Handling financial matters"
#| label: tbl-jorm14

hrs16_func %>%
  QSPtools::checkvar(jorm14, PD545, PD546, PD547) %>%
  gt::gt() %>%
  my_gt() 
  
```

```{r}
#| tbl-cap: "Jorm: Handling everyday arthimetic problems"
#| label: tbl-jorm15

hrs16_func %>%
  QSPtools::checkvar(jorm15, PD548, PD549, PD550) %>%
  gt::gt() %>%
  my_gt()
    
```

```{r}
#| tbl-cap: "Jorm: Using intelligence to understand what's going on"
#| label: tbl-jorm16

hrs16_func %>%
  QSPtools::checkvar(jorm16, PD551, PD552, PD553) %>%
  gt::gt() %>%
  my_gt()
  
```

```{r}
#| tbl-cap: "Jorm Score"
#| label: tbl-jorm-score

foo <- hrs16_func %>%
  left_join(hrshcap, by = c("HHID"="hhid", "PN" = "pn"))

foo %>%
  select(jorm, vs3jormsc) %>%
  gtsummary::tbl_summary(
    type = gtsummary::all_continuous() ~ "continuous2",
    statistic = list(
      gtsummary::all_continuous() ~ c("{mean} ({sd})",
                                      "{median} ({p25}, {p75})",
                                      "{N_nonmiss}"),
      gtsummary::all_categorical() ~ "{n} / {N} ({p}%)"
    ),
  ) %>%
  my_gtsummary()
  
```

```{r}
#| fig-cap: "Distribution of Jorm scores in HRS & HCAP"
#| label: fig-jorm-hist

foo %>%
  select(HHID, PN, jorm, vs3jormsc) %>%
  rename(hrs = jorm) %>%
  rename(hcap = vs3jormsc) %>%
  pivot_longer(cols = c(hrs, hcap), names_to = "source", values_to = "score") %>%
  ggplot(aes(x = score, group = source)) +
    geom_histogram() +
    facet_wrap(~source, ncol=1) +
    hrbrthemes::theme_ipsum()
```

```{r}
#| fig-cap: "Scatterplot of Jorm scores by HRS & HCAP"
#| label: fig-jorm-hrs-hcap

foo %>%
  ggplot(aes(x = jorm, y = vs3jormsc)) +
  geom_point() +
  hrbrthemes::theme_ipsum()
```



**Important note.** The table below shows the labels self-concerns variable based on PD102 (worse than 2 years ago). However the variable as of 2025-06-01 has been defined using both PD101 (fair or po0r memory) AND PD102 (worse than 2 years ago). The labels need to be updated, but the frequencies are still correct (updated). -RNJ

```{r}
#| tbl-cap: "Self-concerns"
#| label: tbl-scd

hrs16_func %>%
  select(self_concerns) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    statistic = list(
      gtsummary::all_continuous() ~ "{mean} ({sd})",
      gtsummary::all_categorical() ~ "{n} / {N} ({p}%)"
    ),
  )


```



