

```{r}
hrs16_cog <- readRDS(fs::path(r_objects_folder, "015_hrs16_cog.rds"))
```

# Recode cognitive items

This section shows the recoding of the cognitive items.

In general, I tried to have higher values indicate correct or better performance.

I recoded "don't know" and "refusal" to incorrect. These could be changed.

If all the component values of a vd variable are NA, then I set the vd variable to NA.

The tables show the recoded variable on the left column, the original variable in the center, and the sample size on the right column.

## Orientation

The number correct of the four orientation to time questions - Month, Day, Year and Day of week.

```{r}
#| tbl-cap: "vdori - Orientation to time - number correct"

hrs16_cog %>%
  mutate(
    PD151 = factor(PD151, levels = c(1, 5, 8, 9),  
                      labels = c("1.  MONTH OK", "5.  MONTH NOT OK", "8.  DK (Don't Know)", 
                                 "9.  RF (Refused)")),
    PD152 = factor(PD152, levels = c(1, 5, 8, 9),  
                      labels = c("1.  DAY OF MONTH OK", "5.  DAY OF MONTH NOT OK", "8.  DK (Don't Know)", 
                                 "9.  RF (Refused)")),
    PD153 = factor(PD153, levels = c(1, 5, 8, 9),  
                      labels = c("1.  YEAR OK", "5.  YEAR NOT OK", "8.  DK (Don't Know)", 
                                 "9.  RF (Refused)")),
    PD154 = factor(PD154, levels = c(1, 5, 8, 9),  
                      labels = c("1.  DAY OK", "5.  DAY NOT OK", "8.  DK (Don't Know)", 
                                 "9.  RF (Refused)"))
    ) %>%
  QSPtools::checkvar(vdori, PD151, PD152, PD153, PD154) %>%
  gt::gt()

```

```{r}
hrs16_cog %>%
  select(vdori) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```

## Count backwards


```{r}
#| tbl-cap: "vdcount: Count backwards"

hrs16_cog %>%
  mutate(
    PD124 = factor(PD124, levels = c(1, 5, 6, 9),  
                      labels = c("1.  CORRECT", "5.  INCORRECT", "6.  WANTS TO START OVER", 
                                 "9.  RF (Refused)")),
    PD129 = factor(PD129, levels = c(1, 5, 9),  
                      labels = c("1.  CORRECT", "5.  INCORRECT", "9.  RF (Refused)"))
    ) %>%
  QSPtools::checkvar(vdcount, PD124, PD129) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdcount) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```

## Object naming

The scissors and cactus naming items have been combined into one item.

```{r}
#| tbl-cap: "vdlfl2: Scissors & Cactus - number correct"

hrs16_cog %>%
  mutate(
    PD155 = factor(PD155, levels = c(1, 5, 8, 9),  
                      labels = c("1.  SCISSORS OR SHEARS ONLY", "5.  NOT CORRECT", "8.  DK (Don't Know)", 
                                 "9.  RF (Refused)")),
    PD156 = factor(PD156, levels = c(1, 5, 8, 9),  
                      labels = c("1.  CACTUS OR NAME OF KIND OF CACTUS", "5.  NOT CORRECT", "8.  DK (Don't Know)",
                                 "9.  RF (Refused)"))
    ) %>%
  QSPtools::checkvar(vdlfl2, PD155, PD156) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdlfl2) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```

## President/Vice-president

The president and vice-president naming items have been combined into one item.

```{r}
#| tbl-cap: "vdlfl3: President and vice-president - number correct"

hrs16_cog %>%
  mutate(
    PD157 = factor(PD157, levels = c(1, 5, 8, 9),  
                      labels = c("1.  LAST NAME CORRECT", "5.  NOT CORRECT", "8.  DK (Don't Know)", 
                                 "9.  RF (Refused)")),
    PD158 = factor(PD158, levels = c(1, 5, 8, 9),  
                      labels = c("1.  LAST NAME CORRECT", "5.  NOT CORRECT", "8.  DK (Don't Know)",
                                 "9.  RF (Refused)"))
    ) %>%
  QSPtools::checkvar(vdlfl3, PD157, PD158) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdlfl3) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```

## Serial 7

The new variable counts the number of correct responses on the serial 7.  Each item was checked so that the difference between consecutive items is 7.  So if a mistake was made early on, there is an opportunity to get the remaining items correct.

```{r}
#| tbl-cap: "PD142-PD146: Serial 7"

hrs16_cog %>%
  mutate(
    rPD142 = factor(rPD142, levels = c(0, 1),  
                      labels = c("0. Incorrect", "1. Correct")),
    rPD143 = factor(rPD143, levels = c(0, 1),  
                      labels = c("0. Incorrect", "1. Correct")),
    rPD144 = factor(rPD144, levels = c(0, 1),  
                      labels = c("0. Incorrect", "1. Correct")),
    rPD145 = factor(rPD145, levels = c(0, 1),  
                      labels = c("0. Incorrect", "1. Correct")),
    rPD146 = factor(rPD146, levels = c(0, 1),  
                      labels = c("0. Incorrect", "1. Correct"))
    ) %>%
  QSPtools::checkvar(vdsevens, rPD142, rPD143, rPD144, rPD145, rPD146) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdsevens) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```

## Animal naming

This variable is the number of correct responses minus the number of errors.  Missing values are set at the median value of errors.

```{r}
#| tbl-cap: "PD196: Animal naming"
QSPtools::checkvar(hrs16_cog, vdlfl1, PD196, PD198) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdlfl1) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```

## Word recall

I found two variables PD174 and PD184 that are the number of words recalled correctly, so we don't need to recreate these scores with PD182M1-PD182M13 and PD183M1-PD183M13.

No recoding was done.

Are we going to use both immediate and delayed recall? Or just delayed recall?

```{r}
#| tbl-cap: "vdwdimm: Immediate word recall"

QSPtools::checkvar(hrs16_cog, vdwdimm, PD174) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdwdimm) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```


```{r}
#| tbl-cap: "vdwddel: Delayed word recall"

QSPtools::checkvar(hrs16_cog, vdwddel, PD184) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdwddel) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```

## Numeracy

Recoded 996, 997, and 999 as missing.

```{r}
#| tbl-cap: "vdexf7: Number series score"

QSPtools::checkvar(hrs16_cog, vdexf7, PNSSCORE) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdexf7) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```

