

```{r}
hrs16_cog <- readRDS(fs::path(r_objects_folder, "015_hrs16_cog.rds"))
```

# Recoding Cognitive Items

This section show preliminary recoding of the cognitive items.

In general, I tried to have higher values indicate correct or better performance.

I recoded "don't know" and "refusal" to incorrect. These could be changed.

If all the component values of a vd variable are NA, then I set the vd variable to NA.

The tables show the recoded variable on the left column, the original variable in the center, and the sample size on the right column.

## Orientation

The number correct of the four orientation to time questions - Month, Day, Year and Day of week.

```{r}
#| tbl-cap: "vdori - Orientation to time - number correct"

QSPtools::checkvar(hrs16_cog, vdori, PD151, PD152, PD153, PD154) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdori) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```

## Count backwards


```{r}
#| tbl-cap: "vdcount: Count backwards"

QSPtools::checkvar(hrs16_cog, vdcount, PD124, PD129) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdcount) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```

## Object naming

The scissors and cactus naming items have been combined into one item.

```{r}
#| tbl-cap: "vdlfl2: Scissors & Cactus - number correct"

QSPtools::checkvar(hrs16_cog, vdlfl2, PD155, PD156) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdlfl2) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```

## President/Vice-president

The president and vice-president naming items have been combined into one item.

```{r}
#| tbl-cap: "vdlfl3: President and vice-president - number correct"

QSPtools::checkvar(hrs16_cog, vdlfl3, PD157, PD158) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdlfl3) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```

## Serial 7

The new variable counts the number of correct responses on the serial 7.  Each item was checked so that the difference between consecutive items is 7.  So if a mistake was made early on, there is an opportunity to get the remaining items correct.

```{r}
#| tbl-cap: "PD142-PD146: Serial 7"
QSPtools::checkvar(hrs16_cog, vdsevens, rPD142, rPD143, rPD144, rPD145, rPD146) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdsevens) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```

## Animal naming

This variable is the number of correct responses minus the number of errors.

```{r}
#| tbl-cap: "PD196: Animal naming"
QSPtools::checkvar(hrs16_cog, vdlfl1, PD196, PD198) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdlfl1) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("r")) ~ "categorical"
  )
```

## Word recall

I found two variables PD174 and PD184 that are the number of words recalled correctly, so we don't need to recreate these scores with PD182M1-PD182M13 and PD183M1-PD183M13.

No recoding was done.

Are we going to use both immediate and delayed recall? Or just delayed recall?

```{r}
#| tbl-cap: "vdwdimm: Immediate word recall"

QSPtools::checkvar(hrs16_cog, vdwdimm, PD174) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdwdimm) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```


```{r}
#| tbl-cap: "vdwddel: Delayed word recall"

QSPtools::checkvar(hrs16_cog, vdwddel, PD184) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdwddel) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```

## Numeracy

Recoded 996, 997, and 999 as missing.

What do these scores mean? They range from 409 - 584.  Do they need to be standardized?

```{r}
#| tbl-cap: "vdexf7: Number series score"

QSPtools::checkvar(hrs16_cog, vdexf7, PNSSCORE) %>%
  gt::gt()
```

```{r}
hrs16_cog %>%
  select(vdexf7) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(
    type = c(starts_with("vd")) ~ "categorical"
  )
```

