
```{r}
hrs16_cog <- readRDS(fs::path(r_objects_folder, "025_hrs16_cog.rds"))
hrs16_cog_norm <- readRDS(fs::path(r_objects_folder, "025_hrs16_cog_norm.rds"))
centering_info <- readRDS(fs::path(r_objects_folder, "025_centering_info.rds"))
foo_lm <- readRDS(fs::path(r_objects_folder, "025_foo_lm.rds"))

goo_lm <- broom::tidy(foo_lm)
```

# Norming the factor score

The HRS sample was filtered to those that were used as the norming sample in HCAP.

The demographic variables were centered to the mean values in the norming sample.  The same centering values were used in the full HRS sample.

These are the centering values used:

```{r}
#| tbl-cap: "Demographic variable means"
#| label: tbl-centering-values
centering_info %>%
  gt::gt() %>%
  gt::fmt_number(decimals = 3) %>%
  my_gt()
```

Summary of the centered demographic variables in the norming sample:

```{r}
#| tbl-cap: "Descriptive statistics of centered demographic variables - HCAP"
#| label: tbl-centered-hcap
hrs16_cog_norm %>%
  select(cfemale, cblack, chisp, cschlyrs) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(statistic = list(
    gtsummary::all_continuous() ~ "{mean} ({sd})",
    gtsummary::all_categorical() ~ "{n} / {N} ({p}%)"
    )
  ) %>%
  my_gtsummary()

```

Summary of the centered demographic variables in the full HRS sample:

```{r}
#| tbl-cap: "Descriptive statistics of centered demographic variables - HRS"
#| label: tbl-centered-hrs
hrs16_cog %>%
  select(cfemale, cblack, chisp, cschlyrs) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_summary(statistic = list(
    gtsummary::all_continuous() ~ "{mean} ({sd})",
    gtsummary::all_categorical() ~ "{n} / {N} ({p}%)"
    )
  ) %>%
  my_gtsummary()

```

Age was modeled with a cubic regression spline. The placement of the knots was determined in the norming sample, using the default percentiles of 5%, 35%, 65%, and 95%. These same knots were used in creating the splines in the full HRS sample.

Theses are the knots used:

```{r}
quantile(hrs16_cog_norm$rage, c(.05, .35, .65, .95))
```

To double check the math, these splines were compared to those from the `rms::rcs()` function.

These are the first 10 rows from the manually created splines.

```{r}
#| tbl-cap: "Manually created regression splines"
#| label: tbl-splines-manual
hrs16_cog_norm %>% 
  slice_head(n=10) %>%
  select(spage1, spage2, spage3) %>%
  gt::gt() %>%
  gt::fmt_number(columns = c(spage2, spage3),
                 decimals = 3) %>%
  my_gt()
```

These are the first 10 rows from the `rms::rcs()` function.

```{r}
#| tbl-cap: "rcs() created regression splines"
#| label: tbl-splines-rcs


# rms::rcs(hrs16_cog_norm$rage, 4)[1:10,]

foo <- rms::rcs(hrs16_cog_norm$rage, 4)
foo <- labelled::remove_attributes(foo, c("class", "name", "label", "assume", "assume.code", "parms", "nonlinear", "colnames"))
foo <- as_tibble(foo)

foo %>% 
  slice_head(n=10) %>%
  gt::gt() %>%
  gt::fmt_number(columns = c(`hrs16_cog_norm'`, `hrs16_cog_norm''`),
                 decimals = 3) %>%
  my_gt()
```

# Factor Score Transformation

The factor score was transformed using a Blom transformation.  The transformation was done in the norming sample.  Cubic regression splines of the factor score were created to predict the Blom score from the factor score.  These cubic regression splines were recreated in the HRS sample to get predicted Blom scores in the HRS sample.


The knots used to create the cubic splines for the Factor score are:

```{r}
quantile(hrs16_cog_norm$F, c(.05, .35, .65, .95), na.rm=TRUE)

```

The regression model is able to predict the blom score well using the factor score cubic splines.

```{r}
#| fig-cap: Original Blom score vs Predicted Blom score in norming sample
#| label: fig-original-predicted-blom

hrs16_cog_norm %>%
  ggplot(aes(x = f_blom, y = Pblom)) +
    geom_abline(intercept = 0, slope = 1, color = "gray") +
    geom_point() +
    geom_smooth(color = "blue") +
    ylim(c(-4, 3)) +
    xlim(c(-4, 3)) +
    hrbrthemes::theme_ipsum() +
    theme(aspect.ratio=1) 


```

The cubic splines using the same knots were recreated in the full HRS sample and were used to predict the blom transformed factor score.

# Normalization of factor scores

The factor scores were normalized by regressing the predicted blom score on the demographic variables described above in the norming sample. Two way interactions between the variables were included.

```{r}
#| tbl-cap: Parameter estimates of the normalization regression in the norming sample
#| label: tbl-normalization-factor-scores

goo_lm %>%
  gt::gt() %>%
  gt::fmt_number(columns = c(estimate, std.error, statistic, p.value),
                 decimals = 3) %>%
  my_gt()

```

## Plots of observed vs model implied blom scores

These are plots of the observed blom scores versus the model implied blom scores that adjusted for demographic characteristics.

```{r}
#| fig-cap: Predicted blom vs normalized blom in norming sample
#| label: fig-pred-normalized-blom-hcap

hrs16_cog_norm %>%
  ggplot(aes(x = Pblom, y = Eblom)) +
  geom_abline(intercept = 0, slope = 1, color = "gray") +
  geom_point() +
  geom_smooth(se = FALSE) +
  ylim(c(-4, 3)) +
  xlim(c(-4, 3)) + 
  hrbrthemes::theme_ipsum() +
  theme(aspect.ratio=1) 
```

The regression model was used to estimate the normalized blom scores in the full HRS sample.

```{r}
#| fig-cap: Predicted blom vs normalized blom in full HRS sample
#| label: fig-pred-normalized-blom-hrs

hrs16_cog %>%
  ggplot(aes(x = Pblom, y = Eblom)) +
  geom_abline(intercept = 0, slope = 1, color = "gray") +
  geom_point() +
  geom_smooth(se=FALSE) +
  ylim(c(-5, 3)) +
  xlim(c(-5, 3)) + 
  hrbrthemes::theme_ipsum() +
  theme(aspect.ratio=1) 
```

# Transforming to T-scores

The normalized blom scores were transformed to a T-score using the following equation:

T= 50 + 10 * ((Predicted blom score - Normalized blom score)/(sd(predicted blom score)*sqrt(1-r2)))

The sd of the predicted blom score in the norming sample is: `r sd(hrs16_cog_norm$Pblom, na.rm=TRUE) %>% round(3)`.

The adjusted model r2 is `r broom::glance(foo_lm) %>% pull(adj.r.squared) %>% round(3)` .

```{r}
#| tbl-cap: Summary statistics of the factor score variables in the norming sample
#| label: tbl-tscores-hcap

hrs16_cog_norm %>%
  select(Pblom, Eblom, TF) %>%
  gtsummary::tbl_summary(
    statistic = list(
    gtsummary::all_continuous() ~ "{mean} ({sd})",
    gtsummary::all_categorical() ~ "{n} / {N} ({p}%)"
    )
  ) %>%
  my_gtsummary()
```

```{r}
#| fig-cap: Predicted blom score vs T-score in norming sample
#| label: fig-pred-tscore-hcap

hrs16_cog_norm %>%
  ggplot(aes(x = Pblom, y = TF)) +
  geom_point() +
  geom_smooth(se=FALSE) + 
  hrbrthemes::theme_ipsum()
```


```{r}
#| tbl-cap: Summary statistics of the factor score variables in the full HRS sample
#| label: tbl-tscores-hrs

hrs16_cog %>%
  select(Pblom, Eblom, TF) %>%
  gtsummary::tbl_summary(
    statistic = list(
    gtsummary::all_continuous() ~ "{mean} ({sd})",
    gtsummary::all_categorical() ~ "{n} / {N} ({p}%)"
    )
  ) %>%
  my_gtsummary()
```

```{r}
#| fig-cap: Predicted blom score vs T-score in the full HRS sample
#| label: fig-pred-tscore-hrs

hrs16_cog %>%
  ggplot(aes(x = Pblom, y = TF)) +
  geom_point() +
  geom_smooth(se=FALSE) + 
  hrbrthemes::theme_ipsum()
```
