
```{r}
tracker_demo <- readRDS(here::here("R_objects", "A0_010_tracker_demo.rds"))

```

## Sample selection variables

This script reads the tracker demographic file and two HCAP auxiliary files, joins them together, and derives a set of sample membership indicators, an age eligibility flag, a unique respondent ID, and a recoded survey cluster variable for use in Mplus. The output is saved as `A0_010_tracker_demo.rds`.


#### `inHRS_16` — In HRS 2016 Core Interview

A binary indicator for whether the respondent completed a core HRS 2016 interview, based on `PIWTYPE`:

| `PIWTYPE` | `inHRS_16` |
|---|---|
| 1 (core interview) | 1 |
| Any other value | 0 |

```{r}
#| tbl-cap: "In HRS 2016 indicator"
QSPtools::checkvar(tracker_demo, inHRS_16, PIWTYPE) %>%
  gt::gt()
```

### `nonzeroweight` — Positive Sampling Weight

A binary indicator for whether the respondent has a positive HRS person-level weight (`PWGTR`):

| Condition | `nonzeroweight` |
|---|---|
| `PWGTR` is missing | 0 |
| `PWGTR > 0` | 1 |
| `PWGTR == 0` | 0 |


### `inHCAP` — In HCAP 2016

A binary indicator for HCAP participation, based on `HCAP16RESP`:

| `HCAP16RESP` | `inHCAP` | Notes |
|---|---|---|
| 1 | 1 | Informant only |
| 2 | 1 | Respondent only |
| 3 | 1 | Both respondent and informant |
| 0 or other | 0 | No participation |

```{r}
#| tbl-cap: "In HCAP 2016 indicator"
QSPtools::checkvar(tracker_demo, inHCAP, HCAP16RESP) %>%
  gt::gt()
```


### `inSamplingFrame` — Weight and HCAP Membership Cross-Classification

A four-category labelled variable summarizing the cross-classification of weight status and HCAP participation:

| Value | Label | Condition |
|---|---|---|
| 1 | In HCAP, has weight | `nonzeroweight == 1` AND `inHCAP == 1` |
| 2 | In HCAP, no weight | `nonzeroweight == 0` AND `inHCAP == 1` |
| 3 | Has weight, not in HCAP | `nonzeroweight == 1` AND `inHCAP == 0` |
| 4 | No weight, not in HCAP | `nonzeroweight == 0` AND `inHCAP == 0` |

```{r}
#| tbl-cap: "In the sampling frame indicator"
QSPtools::checkvar(tracker_demo, inSamplingFrame, nonzeroweight, inHCAP) %>%
  gt::gt()
```

### `age65up` — Age 65 or Older

A binary eligibility indicator derived from self-reported age, with `PA019` (Section A age) taking priority over `PAGE` (tracker age):

| Condition | `age65up` |
|---|---|
| `PA019 >= 65` | 1 |
| `PA019 < 65` | 0 |
| `PA019` missing AND `PAGE >= 65` | 1 |
| `PA019` missing AND `PAGE < 65` | 0 |
| `PA019` missing and `PAGE` missing | `NA` |

```{r}
#| tbl-cap: "Age 65+ indicator"
QSPtools::checkvar(tracker_demo, age65up, PA019, PAGE) %>%
  gt::gt()
```

### `id` — Unique Respondent Identifier

A string identifier formed by concatenating `HHID` and `PN` with `str_c()`. Used as a row-level ID in Mplus analyses.


### `SECU_mplus` — Survey Cluster Variable for Mplus

A recoded cluster variable computed as:

$$
\text{SECU\_mplus} = 10 \times \text{STRATUM} + \text{SECU}
$$

This encoding creates a unique numeric identifier for each SECU-within-STRATUM combination in a format compatible with Mplus survey design specification.








