
```{r}
tracker_demo <- readRDS(here::here("R_objects", "A0_010_tracker_demo.rds"))

```

# 3. Sample identification

This section describes how the tracker demographic file and two HCAP auxiliary files (`hc16hp_r`, `hcap_validation`) are joined to `tracker_demo` and several derived variables are created.  The created variables are a set of sample membership indicators, age eligibility flags, a unique respondent ID, and a recoded survey cluster variable for use in Mplus. Indicators are created for all four HRS waves (2016, 2018, 2020, 2022). 

## Input Data

| Object | Source File | Description |
|---|---|---|
| `tracker_demo` | `A0_009_tracker_demo.rds` | Demographics, weights, survey design variables |
| `hc16hp_r` | `A0_009_hc16hp_r.rds` | HCAP 2016 sampling weights |
| `hcap_validation` | `A0_009_hcap_validation.rds` | HCAP validation subsample (N=50) |

Both `hc16hp_r` and `hcap_validation` are joined to `tracker_demo` by `HHID` and `PN`.

The output is saved as `A0_010_tracker_demo.rds`.


### `inHRS_16` through `inHRS_22` — In HRS Core Interview

Binary indicators for whether the respondent completed a core HRS interview in each wave, based on the wave-specific interview type variable (`xIWTYPE`):

| Condition | Value |
|---|---|
| `xIWTYPE == 1` (core interview) | 1 |
| Any other value | 0 |

```{r}
#| tbl-cap: "In HRS 2016 indicator"
QSPtools::checkvar(tracker_demo, inHRS_16, PIWTYPE) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "In HRS 2018 indicator"
QSPtools::checkvar(tracker_demo, inHRS_18, QIWTYPE) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "In HRS 2020 indicator"
QSPtools::checkvar(tracker_demo, inHRS_20, RIWTYPE) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "In HRS 2022 indicator"
QSPtools::checkvar(tracker_demo, inHRS_22, SIWTYPE) %>%
  gt::gt()
```

### `nonzeroweight16` through `nonzeroweight22` — Positive Sampling Weight

Binary indicators for whether the respondent has a positive HRS person-level weight in each wave (`PWGTR`, `QWGTR`, `RWGTR`, `SWGTR`):

| Condition | Value |
|---|---|
| Weight is missing | 0 |
| Weight > 0 | 1 |
| Weight == 0 | 0 |


### `inHCAP` — In HCAP 2016

A binary indicator for HCAP participation, based on `HCAP16RESP`:

| `HCAP16RESP` | `inHCAP` | Notes |
|---|---|---|
| 1 | 1 | Informant only |
| 2 | 1 | Respondent only |
| 3 | 1 | Both respondent and informant |
| 0 or other | 0 | No participation |

```{r}
#| tbl-cap: "In HCAP 2016 indicator"
QSPtools::checkvar(tracker_demo, inHCAP, HCAP16RESP) %>%
  gt::gt()
```


### `inSamplingFrame` — Weight and HCAP Membership Cross-Classification

A four-category labelled variable summarizing the cross-classification of 2016 weight status and HCAP participation:

| Value | Label | Condition |
|---|---|---|
| 1 | In HCAP, has weight | `nonzeroweight16 == 1` AND `inHCAP == 1` |
| 2 | In HCAP, no weight | `nonzeroweight16 == 0` AND `inHCAP == 1` |
| 3 | Has weight, not in HCAP | `nonzeroweight16 == 1` AND `inHCAP == 0` |
| 4 | No weight, not in HCAP | `nonzeroweight16 == 0` AND `inHCAP == 0` |

```{r}
#| tbl-cap: "In the sampling frame indicator"
QSPtools::checkvar(tracker_demo, inSamplingFrame, nonzeroweight16, inHCAP) %>%
  gt::gt()
```

### `age65up16` through `age65up22` — Age 65 or Older

Binary age eligibility indicators for each wave, derived from self-reported age. The Section A age variable (`xA019`) takes priority over the tracker age variable (`xAGE`):

| Condition | Value |
|---|---|
| `xA019 >= 65` | 1 |
| `xA019 < 65` | 0 |
| `xA019` missing AND `xAGE >= 65` | 1 |
| `xA019` missing AND `xAGE < 65` | 0 |
| Both `xA019` and `xAGE` missing | `NA` |

```{r}
#| tbl-cap: "2016: Age 65+ indicator"
QSPtools::checkvar(tracker_demo, age65up16, PA019, PAGE) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2018: Age 65+ indicator"
QSPtools::checkvar(tracker_demo, age65up18, QA019, QAGE) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2020: Age 65+ indicator"
QSPtools::checkvar(tracker_demo, age65up20, RA019, RAGE) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2022: Age 65+ indicator"
QSPtools::checkvar(tracker_demo, age65up22, SA019, SAGE) %>%
  gt::gt()
```

### `id` — Unique Respondent Identifier

A string identifier formed by concatenating `HHID` and `PN` with `str_c()`. Used as a row-level ID in Mplus analyses.


### `SECU_mplus` — Survey Cluster Variable for Mplus

A recoded cluster variable computed as:

$$
\text{SECU\_mplus} = 10 \times \text{STRATUM} + \text{SECU}
$$

This encoding creates a unique numeric identifier for each SECU-within-STRATUM combination in a format compatible with Mplus survey design specification.


### Output Data

| Object | Output File | Key Content |
|---|---|---|
| `tracker_demo` | `A0_010_tracker_demo.rds` | tracker_demo with all derived sample indicators appended |






