---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
hrs16_cog <- readRDS(here::here("R_objects", "A0_015_hrs16_cog.rds"))
```

## Cognitive Items

This script reads the raw HRS 2016 cognitive items `A0_009A0_009_hrs16_cog.rds` and performs two operations:

1. **Recodes individual cognitive test items** into binary correct/incorrect indicators or cleaned continuous scores
2. **Constructs domain-level summary scores** by summing or combining item-level variables

The output is saved as `A0_015_hrs16_cog.rds`.


### 1. Item-Level Recoding

#### Orientation to Time (`rPD151`–`rPD154`)

Four date/time orientation items are each recoded to binary correct (1) / incorrect (0). Responses of don't know (8) and refuse (9) are scored as **incorrect (0)**.

| Variable | Item | Source |
|---|---|---|
| `rPD151` | Month | PD151 |
| `rPD152` | Day | PD152 |
| `rPD153` | Year | PD153 |
| `rPD154` | Day of week | PD154 |

```{r}
#| tbl-cap: "Month"
hrs16_cog %>%
  QSPtools::checkvar(rPD151, PD151) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "Day"
hrs16_cog %>%
  QSPtools::checkvar(rPD152, PD152) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "Year"
hrs16_cog %>%
  QSPtools::checkvar(rPD153, PD153) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "Day of week"
hrs16_cog %>%
  QSPtools::checkvar(rPD154, PD154) %>%
  gt::gt()
```

#### Object & Person Naming (`rPD155`–`rPD158`)

Four naming items recoded to binary correct (1) / incorrect (0). Don't know and refuse are scored as **incorrect (0)**.

| Variable | Item | Source |
|---|---|---|
| `rPD155` | Scissors (object naming) | PD155 |
| `rPD156` | Cactus (object naming) | PD156 |
| `rPD157` | President (person naming) | PD157 |
| `rPD158` | Vice-president (person naming) | PD158 |

```{r}
#| tbl-cap: "Object naming: Scissors"
hrs16_cog %>%
  QSPtools::checkvar(rPD155, PD155) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "Object naming: Cactus"
hrs16_cog %>%
  QSPtools::checkvar(rPD156, PD156) %>%
  gt::gt()
```

#### Serial 7s Subtraction (`rPD142`–`rPD146`)

Five sequential subtraction items. Each is scored correct (1) if the respondent's answer equals the correct value for that step; don't know (998), refuse (999), and any other non-missing, non-correct values are scored **incorrect (0)**.

Correctness for steps 2–5 is evaluated **relative to the prior response** (i.e., the respondent's previous answer minus 7), not the expected absolute value — preserving credit for internally consistent sequences following an initial error.

| Variable | Item | Correct Answer Rule |
|---|---|---|
| `rPD142` | Serial 7's — 1 | PD142 == 93 |
| `rPD143` | Serial 7's — 2 | PD143 == PD142 − 7 |
| `rPD144` | Serial 7's — 3 | PD144 == PD143 − 7 |
| `rPD145` | Serial 7's — 4 | PD145 == PD144 − 7 |
| `rPD146` | Serial 7's — 5 | PD146 == PD145 − 7 |


```{r}
#| tbl-cap: "Serial sevens"
hrs16_cog %>%
  QSPtools::checkvar(vdsevens, rPD142, rPD143, rPD144, rPD145, rPD146) %>%
  gt::gt()
```

#### Animal Naming Errors (`rPD198`)

`PD198` records the number of animal naming errors. Two special cases are handled before the score is used:

- `PD198 == 98` (coded as missing in HRS) → recoded to **1** (the median error value)
- `NA` → recoded to **1** (median imputation)

All other values are retained as-is.


#### Count Backwards from 20 (`vdcount`)

A binary indicator for whether the respondent correctly counted backwards from 20, allowing for two tries:

| Condition | `vdcount` |
|---|---|
| First try correct (`PD124 == 1`) OR second try correct (`PD129 == 1`) | 1 |
| First try incorrect (`PD124 == 5`) or wants to restart (6) or refuses (9) | 0 |
| Second try incorrect (`PD129 == 5`) or refuses (9) — after failing first try | 1 |

> Note: A failed second try is scored as **1** (partial credit), consistent with the HRS scoring convention for this item.

```{r}
#| tbl-cap: "Count backwards from twenty"
hrs16_cog %>%
  QSPtools::checkvar(vdcount, PD124, PD129) %>%
  gt::gt()
```

#### Word recall

Word recall items are carried forward with minimal transformation:

| Variable | Label | Source | Notes |
|---|---|---|---|
| `vdwdimm` | Word recall — Immediate | PD174 | Direct assignment |
| `vdwddel` | Word recall — Delayed | PD184 | Direct assignment |


#### Number series

Number series item is carried forward with minimal transformation:

| Variable | Label | Source | Notes |
|---|---|---|---|
| `vdexf7` | Number series | PNSSCORE | Codes 996, 997, 999 set to `NA` |




### 2. Domain-Level Summary Scores

All summaries are row-wise. If **all** component items are missing, the score is set to `NA` rather than 0.

| Score | Label | Components | Range |
|---|---|---|---|
| `vdori` | Orientation to Time | `rPD151`–`rPD154` | 0–4 |
| `vdlfl2` | Object naming | `rPD155`, `rPD156` | 0–2 |
| `vdlfl3` | Person naming | `rPD157`, `rPD158` | 0–2 |
| `vdsevens` | Serial 7's — number correct | `rPD142`–`rPD146` | 0–5 |
| `vdlfl1` | Animal naming (correct − errors) | `PD196 − rPD198` | ≥ 0 (floored at 0) |

`vdlfl1` is floored at 0: if the number of errors exceeds the number of correct animals, the score is set to 0 rather than going negative.



