
```{r}
hrs16_cog <- readRDS(here::here("R_objects", "A0_015_hrs16_cog.rds"))
hrs18_cog <- readRDS(here::here("R_objects", "A0_015_hrs18_cog.rds"))
hrs20_cog <- readRDS(here::here("R_objects", "A0_015_hrs20_cog.rds"))
hrs22_cog <- readRDS(here::here("R_objects", "A0_015_hrs22_cog.rds"))
```

## Cognitive Items

This section describes the raw HRS cognitive items for all four waves (2016, 2018, 2020, 2022) and performs two recoding operations. 

1. **Recodes individual cognitive test items** into binary correct/incorrect indicators or cleaned continuous scores
2. **Constructs domain-level summary scores** by summing or combining item-level variables

> **Note:** The 2018 wave (`hrs18_cog`, wave prefix Q) does **not** include the Number Series item (`QNSSCORE` / `rQvdexf7`), as it was not administered in that wave.

> **Note:** As of 2026-05-12, the recoding logic was refactored into a function to support consistent application across waves:


### Input Data

| Object | Source File | Description |
|---|---|---|
| `hrs16_cog` | `A0_009_hrs16_cog.rds` | HRS 2016 raw cognitive items |
| `hrs18_cog` | `A0_009_hrs18_cog.rds` | HRS 2018 raw cognitive items |
| `hrs20_cog` | `A0_009_hrs20_cog.rds` | HRS 2020 raw cognitive items |
| `hrs22_cog` | `A0_009_hrs22_cog.rds` | HRS 2022 raw cognitive items |

The output is saved as `A0_015_hrs{wave}_cog.rds` for each wave.





### 1. Item-Level Recoding

Wave prefixes follow the HRS convention: P = 2016, Q = 2018, R = 2020, S = 2022.

#### Orientation to Time (`r{x}D151`–`r{x}D154`)

Four date/time orientation items are each recoded to binary correct (1) / incorrect (0). Responses of don't know (8) and refuse (9) are scored as **incorrect (0)**.

| Variable | Item | Source |
|---|---|---|
| `r{x}D151` | Month       | `xD151` |
| `r{x}D152` | Day         | `xD152` |
| `r{x}D153` | Year        | `xD153` |
| `r{x}D154` | Day of week | `xD154` |

| Raw Code | Recoded Value |
|----------|---------------|
| 1        | 1 (Correct)   |
| 5        | 0 (Incorrect) |
| 8        | 0 (Incorrect) |
| 9        | 0 (Incorrect) |

```{r}
#| tbl-cap: "2016: Month"
hrs16_cog %>%
  QSPtools::checkvar(rPD151, PD151) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2016: Day"
hrs16_cog %>%
  QSPtools::checkvar(rPD152, PD152) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2016: Year"
hrs16_cog %>%
  QSPtools::checkvar(rPD153, PD153) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2016: Day of week"
hrs16_cog %>%
  QSPtools::checkvar(rPD154, PD154) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2018: Month"
hrs18_cog %>%
  QSPtools::checkvar(rQD151, QD151) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2018: Day"
hrs18_cog %>%
  QSPtools::checkvar(rQD152, QD152) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2018: Year"
hrs18_cog %>%
  QSPtools::checkvar(rQD153, QD153) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2018: Day of week"
hrs18_cog %>%
  QSPtools::checkvar(rQD154, QD154) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2020: Month"
hrs20_cog %>%
  QSPtools::checkvar(rRD151, RD151) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2020: Day"
hrs20_cog %>%
  QSPtools::checkvar(rRD152, RD152) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2020: Year"
hrs20_cog %>%
  QSPtools::checkvar(rRD153, RD153) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2020: Day of week"
hrs20_cog %>%
  QSPtools::checkvar(rRD154, RD154) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2022: Month"
hrs22_cog %>%
  QSPtools::checkvar(rSD151, SD151) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2022: Day"
hrs22_cog %>%
  QSPtools::checkvar(rSD152, SD152) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2022: Year"
hrs22_cog %>%
  QSPtools::checkvar(rSD153, SD153) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2022: Day of week"
hrs22_cog %>%
  QSPtools::checkvar(rSD154, SD154) %>%
  gt::gt()
```

#### Object & Person Naming  (`r{x}D155`–`r{x}D158`)

Four naming items recoded to binary correct (1) / incorrect (0). Don't know and refuse are scored as **incorrect (0)**. The same recoding scheme as orientation items applies.

| Variable | Item | Source |
|---|---|---|
| `r{x}D155` | Scissors (object naming)       | `xD155` |
| `r{x}D156` | Cactus (object naming)         | `xD156` |
| `r{x}D157` | President (person naming)      | `xD157` |
| `r{x}D158` | Vice-president (person naming) | `xD158` |


```{r}
#| tbl-cap: "2016: Object naming: Scissors"
hrs16_cog %>%
  QSPtools::checkvar(rPD155, PD155) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2016: Object naming: Cactus"
hrs16_cog %>%
  QSPtools::checkvar(rPD156, PD156) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2016: Person naming: President"
hrs16_cog %>%
  QSPtools::checkvar(rPD157, PD157) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2016: Person naming: Vice-president"
hrs16_cog %>%
  QSPtools::checkvar(rPD158, PD158) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2018: Object naming: Scissors"
hrs18_cog %>%
  QSPtools::checkvar(rQD155, QD155) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2018: Object naming: Cactus"
hrs18_cog %>%
  QSPtools::checkvar(rQD156, QD156) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2018: Person naming: President"
hrs18_cog %>%
  QSPtools::checkvar(rQD157, QD157) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2018: Person naming: Vice-president"
hrs18_cog %>%
  QSPtools::checkvar(rQD158, QD158) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2020: Object naming: Scissors"
hrs20_cog %>%
  QSPtools::checkvar(rRD155, RD155) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2020: Object naming: Cactus"
hrs20_cog %>%
  QSPtools::checkvar(rRD156, RD156) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2020: Person naming: President"
hrs20_cog %>%
  QSPtools::checkvar(rRD157, RD157) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2020: Person naming: Vice-president"
hrs20_cog %>%
  QSPtools::checkvar(rRD158, RD158) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2022: Object naming: Scissors"
hrs22_cog %>%
  QSPtools::checkvar(rSD155, SD155) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2022: Object naming: Cactus"
hrs22_cog %>%
  QSPtools::checkvar(rSD156, SD156) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2022: Person naming: President"
hrs22_cog %>%
  QSPtools::checkvar(rSD157, SD157) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2022: Person naming: Vice-president"
hrs22_cog %>%
  QSPtools::checkvar(rSD158, SD158) %>%
  gt::gt()
```

#### Serial 7s Subtraction (`r{x}D142`–`r{x}D146`)

Five sequential subtraction items. Each is scored correct (1) if the respondent's answer equals the correct value for that step; don't know (998), refuse (999), and any other non-missing, non-correct values are scored **incorrect (0)**.

Correctness for steps 2–5 is evaluated **relative to the prior response** (i.e., the respondent's previous answer minus 7), not the expected absolute value — preserving credit for internally consistent sequences following an initial error.

| Variable | Item | Correct Answer Rule |
|---|---|---|
| `r{x}D142` | Serial 7's — 1 | `xD142 == 93` |
| `r{x}D143` | Serial 7's — 2 | `xD143 == xD142 − 7` |
| `r{x}D144` | Serial 7's — 3 | `xD144 == xD143 − 7` |
| `r{x}D145` | Serial 7's — 4 | `xD145 == xD144 − 7` |
| `r{x}D146` | Serial 7's — 5 | `xD146 == xD145 − 7` |



```{r}
#| tbl-cap: "2016: Serial sevens"
hrs16_cog %>%
  QSPtools::checkvar(rPvdsevens, rPD142, rPD143, rPD144, rPD145, rPD146) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2018: Serial sevens"
hrs18_cog %>%
  QSPtools::checkvar(rQvdsevens, rQD142, rQD143, rQD144, rQD145, rQD146) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2020: Serial sevens"
hrs20_cog %>%
  QSPtools::checkvar(rRvdsevens, rRD142, rRD143, rRD144, rRD145, rRD146) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2022: Serial sevens"
hrs22_cog %>%
  QSPtools::checkvar(rSvdsevens, rSD142, rSD143, rSD144, rSD145, rSD146) %>%
  gt::gt()
```

#### Animal Naming Errors `r{x}D198`)

`xD198` records the number of animal naming errors. Two special cases are handled before the score is used:

- `xD198 == 98` (coded as missing in HRS) → recoded to **1** (the median error value)
- `NA` → recoded to **1** (median imputation)

All other values are retained as-is.


#### Count Backwards from 20 (`r{x}vdcount`)

A binary indicator for whether the respondent correctly counted backwards from 20, allowing for two tries:

| Condition | `r{x}vdcount` |
|---|---|
| First try correct (`xD124 == 1`) OR second try correct (`xD129 == 1`) | 1 |
| First try incorrect (`xD124 == 5`), wants to restart (6), or refuses (9) | 0 |
| Second try incorrect (`xD129 == 5`) or refuses (9) | 0 |

> Note: A failed second try is scored as **1** (partial credit), consistent with the HRS scoring convention for this item.

```{r}
#| tbl-cap: "2016: Count backwards from twenty"
hrs16_cog %>%
  QSPtools::checkvar(rPvdcount, PD124, PD129) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2018: Count backwards from twenty"
hrs18_cog %>%
  QSPtools::checkvar(rQvdcount, QD124, QD129) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2020: Count backwards from twenty"
hrs20_cog %>%
  QSPtools::checkvar(rRvdcount, RD124, RD129) %>%
  gt::gt()
```

```{r}
#| tbl-cap: "2022: Count backwards from twenty"
hrs22_cog %>%
  QSPtools::checkvar(rSvdcount, SD124, SD129) %>%
  gt::gt()
```

#### Word recall (`r{x}vdwdimm`, `r{x}vdwddel`)

Word recall items are carried forward with direct assignment:

| Variable | Label | Source |
|---|---|---|
| `r{x}vdwdimm` | Word recall: Immediate | `xD174` |
| `r{x}vdwddel` | Word recall: Delayed   | `xD184` |


#### Number series (`r{x}vdexf7`)

Available for 2016, 2020, and 2022 only — **not administered in 2018**. Codes 996, 997, and 999 are set to `NA`; all other non-missing values are retained.


| Variable | Label | Source | Notes |
|---|---|---|---|
| `r{x}vdexf7` | Number series | `xNSSCORE` | Codes 996, 997, 999 set to `NA` |



### 2. Domain-Level Summary Scores

All summaries are row-wise. If **all** component items are missing, the score is set to `NA` rather than 0. The same construction applies across all four waves.

| Score | Label | Components | Range |
|---|---|---|---|
| `r{x}vdori`    | Orientation to Time          | `r{x}D151`–`r{x}D154`                    | 0–4  |
| `r{x}vdlfl2`   | Object naming                | `r{x}D155`, `r{x}D156`                   | 0–2  |
| `r{x}vdlfl3`   | Person naming                | `r{x}D157`, `r{x}D158`                   | 0–2  |
| `r{x}vdsevens` | Serial 7's — number correct  | `r{x}D142`–`r{x}D146`                    | 0–5  |
| `r{x}vdlfl1`   | Animal naming (correct − errors) | `xD196 − r{x}D198`                   | ≥ 0 (floored at 0) |

`r{x}vdlfl1` is floored at 0: if the number of errors exceeds the number of correct animals, the score is set to 0 rather than going negative.


### Output Data

| Object | Output File | Key Content |
|---|---|---|
| `hrs16_cog`       | `A0_015_hrs16_cog.rds`       | 2016 recoded cognitive items and domain scores |
| `hrs16_cog_notes` | `A0_015_hrs16_cog_notes.rds` | Notes table documenting domain score construction |
| `hrs18_cog`       | `A0_015_hrs18_cog.rds`       | 2018 recoded cognitive items and domain scores (no `vdexf7`) |
| `hrs20_cog`       | `A0_015_hrs20_cog.rds`       | 2020 recoded cognitive items and domain scores |
| `hrs22_cog`       | `A0_015_hrs22_cog.rds`       | 2022 recoded cognitive items and domain scores |


