

```{r}
tracker_demo <- readRDS(here::here("R_objects", "A0_011_tracker_demo.rds"))


```


## Demographic variables

This section describes how the tracker demographic file produced in the prior step (`A0_010_tracker_demo.rds`) is used to perform five major operations:

1. **Imputes missing years of schooling** (`SCHLYRSimp`) using predictive mean matching (PMM) via `mice`
2. **Derives descriptive demographic variables** — sex, age categories, race/ethnicity, and educational attainment
3. **Computes restricted cubic spline age terms** for all four waves (2016, 2018, 2020, 2022)
4. **Constructs analysis covariates** (x1–x7 per wave) and all eligible two-way interactions
5. **Centers relevant covariates** using HCAP-weighted means estimated in the HCAP subsample, applied consistently across all four waves


### Input Data

| Object | Source File | Description |
|---|---|---|
| `tracker_demo` | `A0_010_tracker_demo.rds` | Demographics, weights, and sample indicators |

The output is saved as `A0_011_tracker_demo.rds` and `A0_011_covariate_means.rds`.

### 1. Education Imputation (`SCHLYRSimp`)

Years of schooling (`SCHLYRS`) contains missing values coded as `99` and structural zeros in `BIRTHYR`. These are first recoded to `NA`. Five respondents also have manually verified education values (originally recoded in `015-Call-data-CORE-respondent.do`).

| HHID    | PN  | New SCHLYRS value |
|---------|-----|-------------------|
| 134189  | 020 | 14                |
| 502334  | 010 | 12                |
| 502607  | 010 | 10                |
| 905576  | 010 | 16                |
| 906962  | 010 | 12                |

Imputation is then carried out using `mice` with a single imputation (`m = 1`), predictive mean matching (`method = "pmm"`), and seed `2`. Predictors are: `DEGREE`, `GENDER`, `HISPANIC`, `RACE`, and `BIRTHYR`. The imputed values are stored as `SCHLYRSimp`.


### 2. Derived Demographic Variables

#### Age (`page`, `qage`, `rage`, `sage`)

Age variables for each wave are mapped directly from the Section A age variables: `page` = `PA019`, `qage` = `QA019`, `rage` = `RA019`, `sage` = `SA019`.

Age is also grouped into an ordered 6-category factor for each wave (`xage_cat`, `xage_cat_f`):

| Category | Age Range   |
|----------|-------------|
| 1        | 65–69       |
| 2        | 70–74       |
| 3        | 75–79       |
| 4        | 80–84       |
| 5        | 85–89       |
| 6        | 90 and over |

Ages below 65 are set to `NA`. Both a numeric (`xage_cat`) and ordered factor (`xage_cat_f`) version are created for each wave.

#### Sex (`female`)

A binary `female` indicator is derived from `GENDER` (1 = male, 0 = female).

```{r}
QSPtools::checkvar(tracker_demo, female, GENDER) %>%
  gt::gt()
```

#### Race and Ethnicity (`black`, `hisp`, `RaceAndEthnicity`)

Two binary indicators are derived:

-   `black`: Black or African-American and **not** Hispanic
-   `hisp`: Hispanic of any racial background (`HISPANIC` in 1, 2, 3)

```{r}
QSPtools::checkvar(tracker_demo, black, RACE, HISPANIC) %>%
  gt::gt()
```

```{r}
QSPtools::checkvar(tracker_demo, hisp, RACE, HISPANIC) %>%
  gt::gt()
```


A four-category summary variable `RaceAndEthnicity` is also constructed with the following priority logic:

| Priority          | Category                                 | Code |
|-------------------|------------------------------------------|------|
| 1                 | White                                    | 1    |
| 2                 | Black or African-American (Not Hispanic) | 2    |
| 3 (overrides 1–2) | Hispanic (any racial group)              | 3    |
| 4                 | All other racial groups                  | 4    |


```{r}
QSPtools::checkvar(tracker_demo, RaceAndEthnicity, RACE, HISPANIC) %>%
  gt::gt()
```


#### Educational Attainment (`Educational_Attainment`)

A four-category labelled variable is derived from `SCHLYRSimp`:

| Category                 | Years of schooling |
|--------------------------|--------------------|
| \< High school           | \< 12              |
| High school              | = 12               |
| Some college             | 13–13              |
| Education beyond college | ≥ 14               |
| Unknown                  | Missing            |



```{r}
QSPtools::checkvar(tracker_demo, Educational_Attainment, SCHLYRSimp) %>%
  gt::gt()
```

### 3. Age Splines (`spage1`, `spage2`, `spage3`)

Restricted cubic splines (RCS) for age are computed using fixed knots at **70, 78, 86, and 94**, matching the parameterization used in prior HRS/HCAP analyses. The function `ageRCS()` implements the truncated power basis:

-   `spage1`: Linear age term, **centered at 70** (deviation from age 70)
-   `spage2`, `spage3`: Nonlinear spline components (zero at age 70 by construction)

| Wave | Age variable | Output variables |
|------|-------------|-----------------|
| 2016 | `page` | `spage1_16`, `spage2_16`, `spage3_16` |
| 2018 | `qage` | `spage1_18`, `spage2_18`, `spage3_18` |
| 2020 | `rage` | `spage1_20`, `spage2_20`, `spage3_20` |
| 2022 | `sage` | `spage1_22`, `spage2_22`, `spage3_22` |

> The knot placement reflects a hockey-stick-shaped age–cognition relationship in the HCAP normative sample: near-linear decline through most of the age range, with a reversal in the oldest-old due to selective survival. More parameters are therefore allocated to the upper age range.


### 4. Analysis Covariates (x1–x7, per wave)

Seven covariates are defined per wave for use in adjustment and classification models. The wave prefix follows the HRS convention (P=2016, Q=2018, R=2020, S=2022):

| Covariate | Source                  | Notes |
|-----------|-------------------------|-------|
| `x{w}1`   | `spage1_{yr}`           | Centered at age 70; only set for `inHRS_{yr}==1` |
| `x{w}2`   | `spage2_{yr}`           | Zero at age 70 by construction |
| `x{w}3`   | `spage3_{yr}`           | Zero at age 70 by construction |
| `x{w}4`   | `female`                | Centered at HCAP weighted mean |
| `x{w}5`   | `black`                 | Centered at HCAP weighted mean |
| `x{w}6`   | `hisp`                  | Centered at HCAP weighted mean |
| `x{w}7`   | `SCHLYRSimp − 12`       | Centered at 12 years of schooling |

Covariates are only populated for respondents present in the corresponding wave (`inHRS_{yr}==1`); all others are `NA`.

#### Two-Way Interactions

All pairwise interactions among x1–x7 are generated **except**: 

- Interactions among x1, x2, and x3 (spline components with one another)
- The `x{w}5 × x{w}6` interaction (black × hisp), as no respondents in the sample are both Black and Hispanic


Interaction terms are named by concatenation (e.g., `xP1xP4`, `xQ2xQ5`). This pattern is applied identically to all four waves.


#### Centering of Demographic Covariates

`x{w}4` (female), `x{w}5` (black), `x{w}6` (hisp), and their interactions (`x{w}4x{w}5`, `x{w}4x{w}6`) are centered on **HCAP-weighted means** estimated from the `inHCAP == 1` subsample using 2016 survey design variables:

- Cluster: `SECU`
- Strata: `STRATUM`  
- Weight: `HCAP16WGTR`

The five computed means are stored in the named list `covariate_means` and applied consistently to all four waves, so that centering is always relative to the 2016 HCAP reference distribution:

| Object | Stored mean |
|--------|-------------|
| `xP4_mean` | Weighted mean of `female` |
| `xP5_mean` | Weighted mean of `black` |
| `xP6_mean` | Weighted mean of `hisp` |
| `xP4xP5_mean` | Weighted mean of `female × black` |
| `xP4xP6_mean` | Weighted mean of `female × hisp` |

> Although centering means are estimated within the HCAP subsample, the centered covariates are applied to the full `tracker_demo` dataset across all waves.


### Output Data

| Object | Output File | Key Content |
|---|---|---|
| `tracker_demo` | `A0_011_tracker_demo.rds` | Full tracker_demo with imputed education, demographic recodes, age splines, covariates x1–x7, and interactions for all four waves |
| `covariate_means` | `A0_011_covariate_means.rds` | Named list of HCAP-weighted means used for centering, for reproducible application to future datasets |




