

```{r}
tracker_demo <- readRDS(here::here("R_objects", "A0_011_tracker_demo.rds"))


```


## Demographic variables

This script takes the tracker demographic file produced in the prior step (`A0_010_tracker_demo.rds`) and performs three major operations:

1.  **Imputes missing years of schooling** (`SCHLYRSimp`) using predictive mean matching (PMM) via `mice`
2.  **Derives descriptive demographic variables** — sex, age categories, race/ethnicity, and educational attainment
3.  **Constructs analysis covariates** (x1–x7 and all eligible two-way interactions), then **centers** relevant covariates using HCAP-weighted means estimated in the HCAP subsample

The output is saved as `A0_011_tracker_demo.rds`.

### 1. Education Imputation (`SCHLYRSimp`)

Years of schooling (`SCHLYRS`) contains missing values coded as `99` and structural zeros in `BIRTHYR`. These are first recoded to `NA`. Five respondents also have manually verified education values (originally recoded in `015-Call-data-CORE-respondent.do`).

| HHID    | PN  | New SCHLYRS value |
|---------|-----|-------------------|
| 134189  | 020 | 14                |
| 502334  | 010 | 12                |
| 502607  | 010 | 10                |
| 905576  | 010 | 16                |
| 906962  | 010 | 12                |

Imputation is then carried out using `mice` with a single imputation (`m = 1`), predictive mean matching (`method = "pmm"`), and seed `2`. Predictors are: `DEGREE`, `GENDER`, `HISPANIC`, `RACE`, and `BIRTHYR`. The imputed values are stored as `SCHLYRSimp`.


### 2. Derived Demographic Variables

#### Sex (`female`)

A binary `female` indicator is derived from `GENDER` (1 = male, 0 = female).

```{r}
QSPtools::checkvar(tracker_demo, female, GENDER) %>%
  gt::gt()
```

#### Race and Ethnicity (`black`, `hisp`, `RaceAndEthnicity`)

Two binary indicators are derived:

-   `black`: Black or African-American and **not** Hispanic
-   `hisp`: Hispanic of any racial background (`HISPANIC` in 1, 2, 3)

```{r}
QSPtools::checkvar(tracker_demo, black, RACE, HISPANIC) %>%
  gt::gt()
```

```{r}
QSPtools::checkvar(tracker_demo, hisp, RACE, HISPANIC) %>%
  gt::gt()
```


A four-category summary variable `RaceAndEthnicity` is also constructed with the following priority logic:

| Priority          | Category                                 | Code |
|-------------------|------------------------------------------|------|
| 1                 | White                                    | 1    |
| 2                 | Black or African-American (Not Hispanic) | 2    |
| 3 (overrides 1–2) | Hispanic (any racial group)              | 3    |
| 4                 | All other racial groups                  | 4    |


```{r}
QSPtools::checkvar(tracker_demo, RaceAndEthnicity, RACE, HISPANIC) %>%
  gt::gt()
```


#### Age (`rage`, `rage_cat`, `rage_cat_f`)

`rage` maps directly to `PA019` (self-reported age). Age is grouped into an ordered 6-category factor:

| Category | Age Range   |
|----------|-------------|
| 1        | 65–69       |
| 2        | 70–74       |
| 3        | 75–79       |
| 4        | 80–84       |
| 5        | 85–89       |
| 6        | 90 and over |

Ages below 65 are set to `NA`.

#### Educational Attainment (`Educational_Attainment`)

A four-category labelled variable is derived from `SCHLYRSimp`:

| Category                 | Years of schooling |
|--------------------------|--------------------|
| \< High school           | \< 12              |
| High school              | = 12               |
| Some college             | 13–13              |
| Education beyond college | ≥ 14               |
| Unknown                  | Missing            |



```{r}
QSPtools::checkvar(tracker_demo, Educational_Attainment, SCHLYRSimp) %>%
  gt::gt()
```

### 3. Age Splines (`spage1`, `spage2`, `spage3`)

Restricted cubic splines (RCS) for age are computed using fixed knots at **70, 78, 86, and 94**, matching the parameterization used in prior HRS/HCAP analyses. The function `ageRCS()` implements the truncated power basis:

-   `spage1`: Linear age term, **centered at 70** (deviation from age 70)
-   `spage2`, `spage3`: Nonlinear spline components (zero at age 70 by construction)

> The knot placement reflects a hockey-stick-shaped age–cognition relationship in the HCAP normative sample: near-linear decline through most of the age range, with a reversal in the oldest-old due to selective survival. More parameters are therefore allocated to the upper age range.


### 4. Analysis Covariates (x1–x7)

Seven covariates are defined for use in adjustment and classification models:

| Covariate | Source            | Centering                         |
|-----------|-------------------|-----------------------------------|
| `x1`      | `spage1`          | Centered at age 70 (see above)    |
| `x2`      | `spage2`          | Zero at age 70 by construction    |
| `x3`      | `spage3`          | Zero at age 70 by construction    |
| `x4`      | `female`          | Centered at HCAP weighted mean    |
| `x5`      | `black`           | Centered at HCAP weighted mean    |
| `x6`      | `hisp`            | Centered at HCAP weighted mean    |
| `x7`      | `SCHLYRSimp − 12` | Centered at 12 years of schooling |

#### Two-Way Interactions

All pairwise interactions among x1–x7 are generated **except**: - Interactions among x1, x2, and x3 (spline components with one another)

This yields interaction terms named `x1x4`, `x1x5`, ..., `x5x6`, etc.

-   The only two-way interaction deliberately excluded from the covariate set is `black × hisp`, as no respondents in the sample are both Black and Hispanic.

#### Centering of Demographic Covariates

`x4`, `x5`, `x6`, and their interactions (`x4x5`, `x4x6`, `x5x6`) are centered on **HCAP-weighted means** estimated from the `inHCAP == 1` subsample. The survey design uses `SECU` as the cluster ID, `STRATUM` as the strata, and `HCAP16WGTR` as the sampling weight (via `survey::svydesign`).

The six computed means are stored in the named list `covariate_means` for reproducible application to future datasets.

> Although centering means are estimated within the HCAP subsample, the centered covariates are applied to the full `tracker_demo` dataset.




