

```{r}
hcap <- readRDS( fs::path(r_objects_folder, "030_hcap.rds"))
hcap_validation <- readRDS(fs::path(r_objects_folder, "005_hcap_validation.rds"))
langa_weir <- readRDS(fs::path(r_objects_folder, "005_langa_weir.rds"))


```


```{r}
hcap_merged <- hcap %>%
  full_join(hcap_validation, by = c("HHID" = "hhid", "PN" = "pn")) %>%
  left_join(langa_weir, by = c("HHID" = "hhid", "PN" = "pn"))

hcap_merged <- hcap_merged %>%
  mutate(sampling_wt = 1/samplingP)

```

# Comparison of the HRS algorithm to other models

## Comparison of the HRS alorgithm to the HCAP validation sample.  

9 of the 50 validation cases weren't in the HCAP sample

### Unweighted data

```{r}
hcap_merged %>%
  # filter(!is.na(consensuspaneldx)) %>%
  labelled::to_factor() %>%
  gtsummary::tbl_cross(dx_v2, consensuspaneldx) %>%
  my_gtsummary()

```


### Weighted data

The sampling weights were calculated by taking the inverse of the sampling probability (`samplingP`).

```{r}
weighted_tab <- xtabs(sampling_wt ~ dx_v2 + consensuspaneldx, data = hcap_merged) %>%
  round(1) 

kap1 <- psych::cohen.kappa(weighted_tab)

weighted_tab %>%
  as.data.frame.matrix() %>%
  rownames_to_column("dx_v2") %>%
  gt::gt() %>%
  gt::tab_spanner("consensuspaneldx", 2:4) %>%
  my_gt()
```

The weighted kappa is  `r kap1[["confid"]][2,2] %>% round(2)` (`r kap1[["confid"]][2,1] %>% round(2)`. `r kap1[["confid"]][2,3] %>% round(2)`).


## Comparison of HRS algorithm to HCAP algorithm

```{r}
hcap_merged %>%
  labelled::to_factor() %>%
  gtsummary::tbl_cross(dx_v2, vs1hcapdxeap) %>%
  my_gtsummary()

```

```{r}
kap2 <- table(hcap_merged$dx_v2, hcap_merged$vs1hcapdxeap) %>%
  psych::cohen.kappa()
```

The weighted kappa is  `r kap2[["confid"]][2,2] %>% round(2)` (`r kap2[["confid"]][2,1] %>% round(2)`. `r kap2[["confid"]][2,3] %>% round(2)`).

## Comparison of HRS algorithm to Langa-Weir

```{r}
hcap_merged  %>%
  labelled::to_factor() %>%
  gtsummary::tbl_cross(dx_v2, cogfunction2016) %>%
  my_gtsummary()

```

```{r}
kap3 <- table(hcap_merged$dx_v2, hcap_merged$cogfunction2016) %>%
  psych::cohen.kappa()
```

The weighted kappa is  `r kap3[["confid"]][2,2] %>% round(2)` (`r kap3[["confid"]][2,1] %>% round(2)`. `r kap3[["confid"]][2,3] %>% round(2)`).

## Comparison of HCAP validation to Langa-Weir

```{r}
weighted_tab2 <- xtabs(sampling_wt ~ cogfunction2016 + consensuspaneldx, data = hcap_merged) %>%
  round(1) 

kap4 <- psych::cohen.kappa(weighted_tab2)

weighted_tab2 %>%
  as.data.frame.matrix() %>%
  rownames_to_column("hcap_validation") %>%
  gt::gt() %>%
  gt::tab_spanner("consensuspaneldx", 2:4) %>%
  my_gt()

```


The weighted kappa is  `r kap4[["confid"]][2,2] %>% round(2)` (`r kap4[["confid"]][2,1] %>% round(2)`. `r kap4[["confid"]][2,3] %>% round(2)`).




