## Norming Factor Scores

This script builds a **survey-weighted regression model predicting the normed Blom factor score (`Pblom`) from demographic/interaction variables**, using the 2016 HCAP normative sample. It then applies that model to generate **demographically-adjusted expected scores (`Eblom`)** across all three data frames, and converts the residual (observed minus expected, standardized) into a **T-scaled score (`TF`, mean 50, SD 10)** — a common approach for producing norm-referenced cognitive scores adjusted for demographic characteristics.

### Inputs

| Object | Source file | Description |
|----|----|----|
| `hrs16_cog_norm` | `A7_052_hrs16_cog_norm.rds` | Normative HCAP 2016 sample with `Pblom` |
| `hrs16_merged` | `A7_052_hrs16_merged.rds` | Full 2016 HCAP sample with `Pblom` |
| `hrs16_22_long` | `A7_053_hrs16_22_long.rds` | Long-format 2016–2022 panel |

### Step 1: Alias Wave-P ("2016") Variables to Generic Names


- Both `hrs16_cog_norm` and `hrs16_merged` are **cross-sectional 2016 (wave P) data**, so their wave-P-prefixed variables (`xP1`...`xP7` and interaction terms like `xP1xP4`) are copied into generic, wave-agnostic names (`x1`...`x7`, `x1x4`, etc.).
- This lets the same model formula be applied consistently to the 2016 cross-sectional data and the long panel (`hrs16_22_long`), which already stores these variables under the generic `x*` names (from the earlier reshape in A7_053).

### Step 2: Fit the Demographic Model (Survey-Weighted)


- Defines the complex survey design (`SECU` primary sampling unit, `STRATUM` strata, `HCAP16WGTR0` weights) for the normative sample.
- Fits a **survey-weighted GLM** predicting `Pblom` from the seven main-effect demographic terms (`x1`–`x7`) and a set of pairwise interaction terms — fit **only on the normative sample**, mirroring the norming approach used for `Pblom` itself.


### Step 3: Predict Expected ("Eblom") Scores in All Samples


- Uses the fitted model to generate **demographically-expected Blom scores (`Eblom`)** for the normative sample and the full 2016 sample.

For the long panel, predictions are generated on a **filtered subset** (rows with non-missing `x1`, i.e. rows with demographic data available) and then merged back:

- This avoids errors from `augment()`/`predict()` on rows with missing predictors, then reattaches `Eblom` by `id` + `wave`, leaving `NA` for waves without demographic data.

### Step 4: Compute R² and Residual SD for Standardization

- Computes a **pseudo-R²** for the survey GLM as `1 - deviance/null.deviance` (~0.366).
- Computes the **weighted standard deviation** of `Pblom` in the normative sample (~0.973) using `Hmisc::wtd.var()`.
- Combines these to get the **standardized residual SD** (~0.775), which is used as the scaling denominator in the T-score formula below.

### Step 5: Compute T-Scaled Residual Score (`TF`)

Applied identically to all three data frames:

- `TF` is a **T-score** (mean 50, SD 10) representing how far a person's observed Blom score (`Pblom`) deviates from their demographically-expected score (`Eblom`), standardized by the residual SD.
- This produces a cognitive score that is **normed against the HCAP sample and adjusted for demographic characteristics** (age, sex, race/ethnicity, education).

### Outputs

| Object | Saved to |
|----|----|
| `hrs16_merged` | `A7_055_hrs16_merged.rds` |
| `hrs16_cog_norm` | `A7_055_hrs16_cog_norm.rds` |
| `hrs16_22_long` | `A7_055_hrs16_22_long.rds` |
| `foo_glm` (fitted model) | `A7_055_foo_lm.rds` |

### Key Methodological Notes

- The model is fit **only in the normative sample** using **survey weights and design**, consistent with proper norming methodology, then applied out-of-sample to the full 2016 data and the 2016–2022 panel — the same "fit-in-norm, apply-to-all" pattern used for the Blom transform in the prior script (A7_052).
- The wave-P variable aliasing step (Step 1) is necessary because the 2016 cross-sectional frames store demographic/interaction variables with a `xP*` prefix, while the long panel already uses generic `x*` names — aliasing lets one model formula work across all three data structures.
- `TF` is analogous to a **standardized residual T-score**: values above 50 indicate better-than-expected cognitive performance given demographics, values below 50 indicate worse-than-expected performance, in SD-of-10 units.
- Missing-data handling for the long panel (filtering to `!is.na(x1)` before prediction, then left-joining back) preserves all panel rows while avoiding prediction errors on waves lacking demographic covariates.


## Descriptive statistics of factor scores

```{r}
hrs16_merged <- readRDS(here::here("R_objects", "A7_055_hrs16_merged.rds"))
```

```{r}
#| fig-cap: "Distribution of factor scores by HCAP normative sample status"
#| label: fig-f-hcap-normative-hist


F_mean <- hrs16_merged %>%
  filter(!is.na(normexcld)) %>%
  mutate(normexcld_f = factor(normexcld, labels = c("HCAP - Normative sample", "HCAP - Excluded sample"))) %>%
  group_by(normexcld_f) %>%
  summarize(mean = mean(PF, na.rm=TRUE))

hrs16_merged %>%
  filter(!is.na(normexcld)) %>%
  mutate(normexcld_f = factor(normexcld, labels = c("HCAP - Normative sample", "HCAP - Excluded sample"))) %>%
  ggplot(aes(x = PF)) +
  geom_histogram() +
  geom_vline(data = F_mean, aes(xintercept = mean), color = "red") +
  facet_wrap(~normexcld_f, ncol=1) +
  hrbrthemes::theme_ipsum()
```

```{r}
#| tbl-cap: "Descriptive statistics of factor scores by HCAP normative sample status"
#| label: tbl-f-hcap-normative

hrs16_merged %>%
  filter(!is.na(normexcld)) %>%
  mutate(normexcld_f = factor(normexcld, labels = c("HCAP - Normative sample", "HCAP - Excluded sample"))) %>%
  select(PF, normexcld_f) %>%
  gtsummary::tbl_summary(by = normexcld_f,
                         type = all_continuous() ~ "continuous2",
                         statistic = gtsummary::all_continuous2() ~ c(
                           "{mean} ({sd})",
                           "{median} ({p25}, {p75})",
                           "{min}, {max}")
                         ) %>%
  gtsummary::add_overall()



```

```{r}
#| tbl-cap: "Descriptive statistics of adjusted factor scores"
#| label: tbl-blom-desc

hrs16_merged %>%
  mutate(
    x = case_when(inHCAP==0 ~ 1,
                  normexcld==1 ~ 2,
                  normexcld==0 ~ 3),
    x = factor(x, levels = 1:3, labels = c("Not in HCAP", "HCAP - Excluded sample", "HCAP - Normative sample") )
    # normexcld_f = factor(normexcld, labels = c("HCAP - Normative sample", "HCAP - Excluded sample"))
    ) %>%
  select(x, Pblom, Eblom, TF) %>% 
  gtsummary::tbl_summary(by = x,
    type = list(c(Pblom, Eblom, TF ) ~ "continuous2"),
    statistic = gtsummary::all_continuous2() ~ c(
                           "{mean} ({sd})",
                           "{median} ({p25}, {p75})",
                           "{min}, {max}")
  )
```

```{r}
#| fig-cap: "Demographically adjusted factor scores vs original factor score"
#| label: fig-scatter-eblom-plom

# hrs16_cog_norm %>%
#   ggplot(aes(x = Pblom, y = Eblom)) +
#   geom_point() +
#   geom_smooth(se=FALSE) +
#   scale_y_continuous(labelled::get_label_attribute(hrs16_cog_norm$Eblom), limits = c(-4, 3)) +
#   scale_x_continuous(labelled::get_label_attribute(hrs16_cog_norm$Pblom), limits = c(-4, 3)) +
#   hrbrthemes::theme_ipsum() +
#   theme(aspect.ratio=1)


hrs16_merged %>%
  filter(inHCAP==1) %>%
  mutate(
    normexcld_f = factor(normexcld, labels = c("HCAP - Normative sample", "HCAP - Excluded sample"))
    ) %>%
  ggplot(aes(x = Pblom, y = Eblom, color = normexcld_f)) +
  geom_point(alpha=.2) +
  geom_smooth(se=FALSE) +
  scale_y_continuous(labelled::get_label_attribute(hrs16_cog_norm$Eblom), limits = c(-4, 3)) +
  scale_x_continuous(labelled::get_label_attribute(hrs16_cog_norm$Pblom), limits = c(-4, 3)) +
  scale_color_discrete("HCAP sample") +
  hrbrthemes::theme_ipsum() +
  theme(aspect.ratio=1)
```

```{r}
#| fig-cap: "T-scaled factor residual vs original factor score"
#| label: fig-scatter-tf-plom

# hrs16_cog_norm %>%
#   ggplot(aes(x = Pblom, y = TF)) +
#   geom_point() +
#   geom_smooth(se=FALSE) +
#   scale_y_continuous(labelled::get_label_attribute(hrs16_cog_norm$TF)) +
#   scale_x_continuous(labelled::get_label_attribute(hrs16_cog_norm$Pblom), limits = c(-4, 3)) +
#   hrbrthemes::theme_ipsum()


hrs16_merged %>%
  filter(inHCAP==1) %>%
  mutate(
    normexcld_f = factor(normexcld, labels = c("HCAP - Normative sample", "HCAP - Excluded sample"))
    ) %>%
  ggplot(aes(x = Pblom, y = TF, color = normexcld_f)) +
  geom_point(alpha = .2) +
  geom_smooth(se=FALSE) +
  scale_y_continuous(labelled::get_label_attribute(hrs16_cog_norm$TF)) +
  scale_x_continuous(labelled::get_label_attribute(hrs16_cog_norm$Pblom), limits = c(-4, 3)) +
  scale_color_discrete("HCAP sample") +
  hrbrthemes::theme_ipsum()

```
