---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
item_df <- tribble(~item, ~label,
"vdori",  "Orientation to time",
"vdlfl1z",  "Animal naming",
"vdlfl2",  "Scissors & cactus",
"vdlfl3",  "President & vice-president",
"vdcount", "Count backwards from 20",
"vdsevens", "Serial sevens",
"vdwdimmz", "Immediate word recall",
"vdwddelz", "Delayed word recall",
"vdexf7z", "Number series"
)

```

# Mplus models

## Model 1

Sample HRS 2016, age 65+

Weights: HCAP16WGTR, participants not in HCAP get a weight of 0

Estimator: WLSMV N = 9218

Missing cognitive testing N=571

Fit statistics:

```{r}
#| tbl-cap: "Model 1: Fit statistics"
#| label: tbl-model1-fit

mplush5::mplus.print.model.fit.information(here::here("mplus_output", "A7", "model1", "model1.h5")) %>%
  filter(grepl("RMSEA : E", description) |
         grepl("SRMR", description) |
         grepl("CFI", description) ) %>%
  gt::gt() %>%
  gt::fmt_number(
    columns = value,
    decimals = 3
  ) %>%
  gt::cols_label(
    description = "Fit statistic",
    value = "Value"
  ) 

```

Factor loadings:

```{r}
#| tbl-cap: "Model 1: Factor loadings"
#| label: tbl-model1-loadings

mplush5::mplus.print.standardized.model.results(here::here("mplus_output", "A7", "model1", "model1.h5"), "stdyx") %>%
  filter(grepl("F BY", Statement))  %>%
  separate(Statement, into = c("f", "by", "item"), sep = " ") %>%
  mutate(item = str_to_lower(item)) %>%
  left_join(item_df, by = c("item" = "item")) %>%
  select(item, label, Estimate) %>%
  gt::gt() %>%
  gt::fmt_number(
    columns = Estimate,
    decimals = 3
  ) %>%
  gt::cols_label(
    item = "Item",
    label = "Label",
    Estimate = "Std Factor Loading"
  ) 

```

## Model 1a

Sample HRS 2016, age 65+

Weights: HCAP16WGTR, participants not in HCAP get a weight of 0

Estimator: MLR-Probit

N = 9218

Missing cognitive testing N=571

Factor loadings:

```{r}
#| tbl-cap: "Model 1a: Factor loadings"
#| label: tbl-model1a-loadings

mplush5::mplus.print.standardized.model.results(here::here("mplus_output", "A7", "model1a", "model1a.h5"), "stdyx") %>%
  filter(grepl("F BY", Statement))  %>%
  separate(Statement, into = c("f", "by", "item"), sep = " ") %>%
  mutate(item = str_to_lower(item)) %>%
  left_join(item_df, by = c("item" = "item")) %>%
  select(item, label, Estimate) %>%
  gt::gt() %>%
  gt::fmt_number(
    columns = Estimate,
    decimals = 3
  ) %>%
  gt::cols_label(
    item = "Item",
    label = "Label",
    Estimate = "Std Factor Loading"
  ) 

```

## Model 1a fixed

Model with all parameters fixed to those from Model 1a, and mean=0 and variance=1.

Estimator: MLR-Probit

N = 9218,

Missing cognitive testing N=571

Note: Everyone gets a factor score, even if they weren't in HCAP

```{r}
hrs16_merged <- readRDS(here::here("R_objects", "A7_050_hrs16_merged.rds"))

```

```{r}
#| fig-cap: "Distribution of factor scores by HCAP inclusion"
#| label: fig-f-hcap-hist


F_mean <- hrs16_merged %>%
  mutate(inHCAP_f = factor(inHCAP, levels = 0:1, labels = c("Not in HCAP", "In HCAP"))) %>%
  group_by(inHCAP_f) %>%
  summarize(mean = mean(F, na.rm=TRUE))

hrs16_merged %>%
  mutate(inHCAP_f = factor(inHCAP, levels = 0:1, labels = c("Not in HCAP", "In HCAP"))) %>%
  ggplot(aes(x = F)) +
  geom_histogram() +
  geom_vline(data = F_mean, aes(xintercept = mean), color = "red") +
  facet_wrap(~inHCAP_f, ncol=1) +
  hrbrthemes::theme_ipsum()
```
