---
output: html_document
editor_options: 
  chunk_output_type: console
---



# Weighted Comparison Tables

Comparisons that involve the consensus diagnosis use the HRS/HCAP validation-subsample weights computed as `HCAP16WGTR * samplingP^(-1)`.

Comparisons that do not involve the consensus diagnosis use the HCAP population weights `HCAP16WGTR`.

```{r}
hcap_tables <- readRDS(here::here("R_objects", "A7_100_hcap_tables.rds"))
```

### Langa-Weir versus Consensus panel

```{r}
#| tbl-cap: "Langa-Weir diagnosis vs Consensus panel diagnosis, with consensus-panel sample weights"
#| label: tbl-lw-consensus-3
format_weighted_tab(
  hcap_tables$tabs$three_class$lw_consensus,
  row_name = "langa_weir",
  stubhead = "Langa-Weir classification",
  spanner = "Consensus panel diagnosis"
)
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$consensus_standard`
:::

The quadratic weighted kappa is `r round(hcap_tables$summary$lw_consensus_kappa, 3)`.

### Algorithmic diagnosis versus Consensus panel

```{r}
#| tbl-cap: "Algorithmic diagnosis vs Consensus panel diagnosis, with consensus-panel sample weights"
#| label: tbl-algorithmic-consensus-3
hcap_tables$tables$algorithmic_consensus
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$consensus_standard`
:::

The quadratic weighted kappa is `r round(hcap_tables$summary$algorithmic_consensus_kappa, 3)`.

### HRS classification model versus Consensus panel

```{r}
#| tbl-cap: "Consensus panel diagnosis vs HRS classification model, with consensus-panel sample weights"
#| label: tbl-consensus-hrs-3
hcap_tables$tables$consensus_hrs
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$consensus_hrs`
:::

The quadratic weighted kappa is `r round(hcap_tables$summary$consensus_hrs_kappa, 3)`.

### Hudomiet versus Consensus panel

```{r}
#| tbl-cap: "Consensus panel diagnosis vs Hudomiet, with consensus-panel sample weights"
#| label: tbl-consensus-hudomiet-3
hcap_tables$tables$consensus_hudomiet
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$consensus_hudomiet`
:::

The quadratic weighted kappa is `r round(hcap_tables$summary$consensus_hudomiet_kappa, 3)`.

### Langa-Weir versus Algorithmic diagnosis

```{r}
#| tbl-cap: "Langa-Weir diagnosis vs Algorithmic diagnosis, with HCAP weights"
#| label: tbl-algorithmic-lw-3
hcap_tables$tables$algorithmic_lw
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$hcap_standard`
:::

The quadratic weighted kappa is `r round(hcap_tables$summary$algorithmic_lw_kappa, 3)`.


### HRS classification model versus Algorithmic diagnosis

```{r}
#| tbl-cap: "Algorithmic diagnosis vs HRS classification model, with HCAP weights"
#| label: tbl-algorithmic-hrs-3
hcap_tables$tables$algorithmic_hrs
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$hcap_hrs`
:::

The quadratic weighted kappa is `r round(hcap_tables$summary$algorithmic_hrs_kappa, 3)`.

### Hudomiet model versus Algorithmic diagnosis

```{r}
#| tbl-cap: "Algorithmic diagnosis vs Hudomiet, with HCAP weights"
#| label: tbl-algorithmic-hudomiet-3
hcap_tables$tables$algorithmic_hudomiet
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$hcap_hudomiet`
:::

The quadratic weighted kappa is `r round(hcap_tables$summary$algorithmic_hudomiet_kappa, 3)`.


### Langa-Weir versus HRS classification model

```{r}
#| tbl-cap: "Langa-Weir diagnosis vs HRS classification model, with HCAP weights"
#| label: tbl-lw-hrs-3
hcap_tables$tables$lw_hrs
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$hcap_hrs`
:::

The quadratic weighted kappa is `r round(hcap_tables$summary$lw_hrs_kappa, 3)`.

### Langa-Weir versus Hudomiet

```{r}
#| tbl-cap: "Langa-Weir diagnosis vs Hudomiet, with HCAP weights"
#| label: tbl-lw-hudomiet-3
hcap_tables$tables$lw_hudomiet
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$hcap_hudomiet`
:::

The quadratic weighted kappa is `r round(hcap_tables$summary$lw_hudomiet_kappa, 3)`.

### HRS classification model versus Hudomiet

```{r}
#| tbl-cap: "HRS classification model vs Hudomiet, with HCAP weights"
#| label: tbl-hrs-hudomiet-3
hcap_tables$tables$hrs_hudomiet
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$hcap_hudomiet`
:::

The quadratic weighted kappa is `r round(hcap_tables$summary$hrs_hudomiet_kappa, 3)`.


### Agreement Matrix

```{r}
#| tbl-cap: "Pairwise quadratic weighted kappa matrix using consensus-panel weights for consensus comparisons and HCAP weights otherwise"
#| label: tbl-agreement-matrix-3
hcap_tables$tables$agreement_matrix
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$matrix_three`
:::

## Dementia Versus Non-dementia

This section collapses Normal and MCI into a single non-demented category. Comparisons that involve the consensus diagnosis use consensus-panel weights, and comparisons that do not involve the consensus diagnosis use HCAP weights.

### Algorithmic diagnosis versus Consensus panel

```{r}
#| tbl-cap: "Algorithmic dementia classification vs Consensus panel diagnosis, with consensus-panel sample weights"
#| label: tbl-algorithmic-consensus-2
hcap_tables$tables$algorithmic_consensus_binary
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$consensus_standard`
:::

The kappa is `r round(hcap_tables$summary$algorithmic_consensus_binary_kappa, 3)`.

### Langa-Weir versus Consensus panel

```{r}
#| tbl-cap: "Langa-Weir dementia classification vs Consensus panel diagnosis, with consensus-panel sample weights"
#| label: tbl-lw-consensus-2
hcap_tables$tables$lw_consensus_binary
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$consensus_standard`
:::

The kappa is `r round(hcap_tables$summary$lw_consensus_binary_kappa, 3)`.

### HRS classification model versus Consensus panel

```{r}
#| tbl-cap: "Consensus panel dementia classification vs HRS classification model, with consensus-panel sample weights"
#| label: tbl-consensus-hrs-2
hcap_tables$tables$consensus_hrs_binary
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$consensus_hrs`
:::

The kappa is `r round(hcap_tables$summary$consensus_hrs_binary_kappa, 3)`.

### Hudomiet versus Consensus panel

```{r}
#| tbl-cap: "Consensus panel dementia classification vs Hudomiet, with consensus-panel sample weights"
#| label: tbl-consensus-hudomiet-2
hcap_tables$tables$consensus_hudomiet_binary
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$consensus_hudomiet`
:::

The kappa is `r round(hcap_tables$summary$consensus_hudomiet_binary_kappa, 3)`.

### Langa-Weir versus Algorithmic diagnosis

```{r}
#| tbl-cap: "Langa-Weir dementia classification vs Algorithmic diagnosis, with HCAP weights"
#| label: tbl-algorithmic-lw-2
hcap_tables$tables$algorithmic_lw_binary
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$hcap_standard`
:::

The kappa is `r round(hcap_tables$summary$algorithmic_lw_binary_kappa, 3)`.


### HRS classification model versus Algorithmic diagnosis

```{r}
#| tbl-cap: "Algorithmic dementia classification vs HRS classification model, with HCAP weights"
#| label: tbl-algorithmic-hrs-2
hcap_tables$tables$algorithmic_hrs_binary
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$hcap_hrs`
:::

The kappa is `r round(hcap_tables$summary$algorithmic_hrs_binary_kappa, 3)`.

### Hudomiet versus Algorithmic diagnosis

```{r}
#| tbl-cap: "Algorithmic dementia classification vs Hudomiet, with HCAP weights"
#| label: tbl-algorithmic-hudomiet-2
hcap_tables$tables$algorithmic_hudomiet_binary
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$hcap_hrs`
:::

The kappa is `r round(hcap_tables$summary$algorithmic_hudomiet_binary_kappa, 3)`.

### Langa-Weir versus HRS classification model

```{r}
#| tbl-cap: "Langa-Weir dementia classification vs HRS classification model, with HCAP weights"
#| label: tbl-lw-hrs-2
hcap_tables$tables$lw_hrs_binary
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$hcap_hrs`
:::

The kappa is `r round(hcap_tables$summary$lw_hrs_binary_kappa, 3)`.

### Langa-Weir versus Hudomiet

```{r}
#| tbl-cap: "Langa-Weir dementia classification vs Hudomiet, with HCAP weights"
#| label: tbl-lw-hudomiet-2
hcap_tables$tables$lw_hudomiet_binary
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$hcap_hrs`
:::

The kappa is `r round(hcap_tables$summary$lw_hudomiet_binary_kappa, 3)`.

### HRS classification model versus Hudomiet

```{r}
#| tbl-cap: "HRS classification model vs Hudomiet, with HCAP weights"
#| label: tbl-hrs-hudomiet-2
hcap_tables$tables$hrs_hudomiet_binary
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$hcap_hrs`
:::

The kappa is `r round(hcap_tables$summary$hrs_hudomiet_binary_kappa, 3)`.

### Binary Agreement Matrix

```{r}
#| tbl-cap: "Pairwise kappa matrix for dementia versus non-dementia using consensus-panel weights for consensus comparisons and HCAP weights otherwise"
#| label: tbl-agreement-matrix-2
hcap_tables$tables$agreement_matrix_binary
```

::: {custom-style="Footnote text"}
`r hcap_tables$notes$matrix_binary`
:::


