## HRS Classification (A7) vs. HCAP Algorithmic Classification

::: {style="font-size: 0.7em;"}
```{r}
#| tbl-cap: "HRS Classification (A7) vs. HCAP Algorithmic classification — unweighted counts"
#| label: tbl-hrs-algorithmic-unweighted

hcap_tables$data$hcap_sample |>
  dplyr::filter(!is.na(algorithmic), !is.na(hrs)) |>
  dplyr::count(hrs, algorithmic) |>
  tidyr::pivot_wider(names_from = algorithmic, values_from = n, values_fill = 0L) |>
  dplyr::rename(`HRS Classification (A7)` = hrs) |>
  gt::gt(rowname_col = "HRS Classification (A7)") |>
  gt::tab_spanner("HCAP Algorithmic classification", columns = -1) |>
  gt::tab_stubhead("HRS Classification (A7)") |>
  gt::grand_summary_rows(
    fns = list(Total = ~ sum(.)),
    fmt = ~ gt::fmt_integer(.)
  )
```
:::

::: {style="font-size: 0.55em; line-height: 1.2;"}
Unweighted N = `r hcap_tables$data$hcap_sample |> dplyr::filter(!is.na(algorithmic), !is.na(hrs)) |> nrow()` HCAP participants.
Quadratic weighted κ = `r round(hcap_tables$summary$algorithmic_hrs_kappa, 3)` (HCAP population weights).
:::

## <span style="font-size: 0.85em;">HRS Classification (A7) vs. HCAP Algorithmic — Weighted %</span>

::: {style="font-size: 0.7em;"}
```{r}
#| tbl-cap: "HRS Classification (A7) vs. HCAP Algorithmic classification — complex-sample weighted percentages"
#| label: tbl-hrs-algorithmic-weighted-pct

{
  tab <- hcap_tables$tabs$three_class$algorithmic_hrs
  pct <- prop.table(tab) * 100

  pct_df <- as.data.frame.matrix(pct) |>
    tibble::rownames_to_column("Algorithmic") |>
    dplyr::mutate(Total = Normal + MCI + Dementia)

  total_row <- pct_df |>
    dplyr::summarise(
      Algorithmic = "Total",
      dplyr::across(c(Normal, MCI, Dementia, Total), sum)
    )

  dplyr::bind_rows(pct_df, total_row) |>
    gt::gt(rowname_col = "Algorithmic") |>
    gt::tab_spanner("HRS Classification (A7)", columns = c(Normal, MCI, Dementia)) |>
    gt::tab_stubhead("HCAP Algorithmic classification") |>
    gt::fmt_number(decimals = 1)
}
```
:::

::: {style="font-size: 0.55em; line-height: 1.2;"}
`r hcap_tables$notes$hcap_hrs`
Quadratic weighted κ = `r round(hcap_tables$summary$algorithmic_hrs_kappa, 3)`.
:::

## <span style="font-size: 0.6em;">3-way Agreement Matrix (Normal / MCI / Dementia)</span>

```{r}
#| tbl-cap: "Pairwise quadratic weighted kappa — 3-way classification"
#| label: tbl-kappa-3way

hcap_tables$tables$agreement_matrix |>
  gt::cols_label(
    `Algorithmic diagnosis`    = "HCAP Algorithmic classification",
    `HRS classification model` = "HRS Classification (A7)",
    `Hudomiet`                 = "Hudomiet BLVM"
  ) |>
  gt::text_transform(
    locations = gt::cells_stub(),
    fn = function(x) {
      x <- gsub("Algorithmic diagnosis", "HCAP Algorithmic classification", x)
      x <- gsub("HRS classification model", "HRS Classification (A7)", x)
      x <- gsub("^Hudomiet$", "Hudomiet BLVM", x)
      x
    }
  )
```

::: {style="font-size: 0.6em; line-height: 1.2;"}
`r hcap_tables$notes$matrix_three`
:::

## <span style="font-size: 0.6em;">2-way Agreement Matrix (Dementia / Non-dementia)</span>

::: {style="font-size: 0.5em;"}
```{r}
#| tbl-cap: "Pairwise kappa — dementia vs. non-dementia"
#| label: tbl-kappa-2way

hcap_tables$tables$agreement_matrix_binary |>
  gt::cols_label(
    `Algorithmic diagnosis`    = "HCAP Algorithmic classification",
    `HRS classification model` = "HRS Classification (A7)",
    `Hudomiet`                 = "Hudomiet BLVM"
  ) |>
  gt::text_transform(
    locations = gt::cells_stub(),
    fn = function(x) {
      x <- gsub("Algorithmic diagnosis", "HCAP Algorithmic classification", x)
      x <- gsub("HRS classification model", "HRS Classification (A7)", x)
      x <- gsub("^Hudomiet$", "Hudomiet BLVM", x)
      x
    }
  )
```
:::

::: {style="font-size: 0.6em; line-height: 1.2;"}
`r hcap_tables$notes$matrix_binary`
:::
