---
output: html_document
editor_options:
  chunk_output_type: console
---

## Weighted Comparison Tables

Comparisons that involve the consensus diagnosis use the HRS/HCAP validation-subsample weights computed as `HCAP16WGTR * samplingP^(-1)`.

Comparisons that do not involve the consensus diagnosis use the HCAP population weights `HCAP16WGTR`.

```{r}
source(here::here("R", "PMM_110_Comparison_of_Consensus_Langa_Weir.R"))
```

### Langa-Weir versus Consensus panel

```{r}
#| tbl-cap: "Langa-Weir diagnosis vs Consensus panel diagnosis, with consensus-panel sample weights"
format_weighted_tab(
  PMM_110$tabs$three_class$lw_consensus,
  row_name = "langa_weir",
  stubhead = "Langa-Weir classification",
  spanner = "Consensus panel diagnosis"
)
```

::: {custom-style="Footnote text"}
`r PMM_110$notes$consensus_standard`
:::

The quadratic weighted kappa is `r round(PMM_110$summary$lw_consensus_kappa, 3)`.

### Algorithmic diagnosis versus Consensus panel

```{r}
#| tbl-cap: "Algorithmic diagnosis vs Consensus panel diagnosis, with consensus-panel sample weights"
PMM_110$tables$algorithmic_consensus
```

::: {custom-style="Footnote text"}
`r PMM_110$notes$consensus_standard`
:::

The quadratic weighted kappa is `r round(PMM_110$summary$algorithmic_consensus_kappa, 3)`.

### Langa-Weir versus Algorithmic diagnosis

```{r}
#| tbl-cap: "Langa-Weir diagnosis vs Algorithmic diagnosis, with HCAP weights"
PMM_110$tables$algorithmic_lw
```

::: {custom-style="Footnote text"}
`r PMM_110$notes$hcap_standard`
:::

The quadratic weighted kappa is `r round(PMM_110$summary$algorithmic_lw_kappa, 3)`.

### Consensus panel versus Combined model classification

```{r}
#| tbl-cap: "Consensus panel diagnosis vs Combined model classification, with consensus-panel sample weights"
PMM_110$tables$consensus_combined
```

::: {custom-style="Footnote text"}
`r PMM_110$notes$consensus_combined`
:::

The quadratic weighted kappa is `r round(PMM_110$summary$consensus_combined_kappa, 3)`.

### Algorithmic diagnosis versus Combined model classification

```{r}
#| tbl-cap: "Algorithmic diagnosis vs Combined model classification, with HCAP weights"
PMM_110$tables$algorithmic_combined
```

::: {custom-style="Footnote text"}
`r PMM_110$notes$hcap_combined`
:::

The quadratic weighted kappa is `r round(PMM_110$summary$algorithmic_combined_kappa, 3)`.

### Langa-Weir versus Combined model classification

```{r}
#| tbl-cap: "Langa-Weir diagnosis vs Combined model classification, with HCAP weights"
PMM_110$tables$lw_combined
```

::: {custom-style="Footnote text"}
`r PMM_110$notes$hcap_combined`
:::

The quadratic weighted kappa is `r round(PMM_110$summary$lw_combined_kappa, 3)`.

### Agreement Matrix

```{r}
#| tbl-cap: "Pairwise quadratic weighted kappa matrix using consensus-panel weights for consensus comparisons and HCAP weights otherwise"
PMM_110$tables$agreement_matrix
```

::: {custom-style="Footnote text"}
`r PMM_110$notes$matrix_three`
:::

## Dementia Versus Non-dementia

This section collapses Normal and MCI into a single non-demented category. Comparisons that involve the consensus diagnosis use consensus-panel weights, and comparisons that do not involve the consensus diagnosis use HCAP weights.

### Langa-Weir versus Consensus panel

```{r}
#| tbl-cap: "Langa-Weir dementia classification vs Consensus panel diagnosis, with consensus-panel sample weights"
PMM_110$tables$lw_consensus_binary
```

::: {custom-style="Footnote text"}
`r PMM_110$notes$consensus_standard`
:::

The kappa is `r round(PMM_110$summary$lw_consensus_binary_kappa, 3)`.

### Algorithmic diagnosis versus Consensus panel

```{r}
#| tbl-cap: "Algorithmic dementia classification vs Consensus panel diagnosis, with consensus-panel sample weights"
PMM_110$tables$algorithmic_consensus_binary
```

::: {custom-style="Footnote text"}
`r PMM_110$notes$consensus_standard`
:::

The kappa is `r round(PMM_110$summary$algorithmic_consensus_binary_kappa, 3)`.

### Langa-Weir versus Algorithmic diagnosis

```{r}
#| tbl-cap: "Langa-Weir dementia classification vs Algorithmic diagnosis, with HCAP weights"
PMM_110$tables$algorithmic_lw_binary
```

::: {custom-style="Footnote text"}
`r PMM_110$notes$hcap_standard`
:::

The kappa is `r round(PMM_110$summary$algorithmic_lw_binary_kappa, 3)`.

### Consensus panel versus Combined model classification

```{r}
#| tbl-cap: "Consensus panel dementia classification vs Combined model classification, with consensus-panel sample weights"
PMM_110$tables$consensus_combined_binary
```

::: {custom-style="Footnote text"}
`r PMM_110$notes$consensus_combined`
:::

The kappa is `r round(PMM_110$summary$consensus_combined_binary_kappa, 3)`.

### Algorithmic diagnosis versus Combined model classification

```{r}
#| tbl-cap: "Algorithmic dementia classification vs Combined model classification, with HCAP weights"
PMM_110$tables$algorithmic_combined_binary
```

::: {custom-style="Footnote text"}
`r PMM_110$notes$hcap_combined`
:::

The kappa is `r round(PMM_110$summary$algorithmic_combined_binary_kappa, 3)`.

### Langa-Weir versus Combined model classification

```{r}
#| tbl-cap: "Langa-Weir dementia classification vs Combined model classification, with HCAP weights"
PMM_110$tables$lw_combined_binary
```

::: {custom-style="Footnote text"}
`r PMM_110$notes$hcap_combined`
:::

The kappa is `r round(PMM_110$summary$lw_combined_binary_kappa, 3)`.

### Binary Agreement Matrix

```{r}
#| tbl-cap: "Pairwise kappa matrix for dementia versus non-dementia using consensus-panel weights for consensus comparisons and HCAP weights otherwise"
PMM_110$tables$agreement_matrix_binary
```

::: {custom-style="Footnote text"}
`r PMM_110$notes$matrix_binary`
:::
