


## Model - GCP factor

This model combined the neuropsych items into a GCP. It allowed the GCP, individual functional impairment indicators, subjective memory complaints, and informant scores to vary by class.

```{r}
pmm_102b <- MplusAutomation::readModels(here::here("mplus_output", "pmm_102", "pmm_hcap_102b.out"))

PMM_100 <- readRDS(here::here("R_objects", "PMM_100.RDS"))
inhcap <- PMM_100 %>%
  filter(inHCAP==1)

```

### Parameter estimates

These are the parameter estimates to show which were constrained to be equal or allowed to vary by class.

In this model all the factor loading were constrained to be the same across class.

```{r}
#| tbl-cap: "Unstandardized factor loadings by class"
pmm_102b[["parameters"]][["unstandardized"]] %>%
  filter(!grepl("ON", paramHeader)) %>%
  filter(grepl("BY", paramHeader)) %>%
  select(param, est, LatentClass) %>%
  pivot_wider(names_from = LatentClass, values_from = est, names_prefix = "class_") %>%
  gt::gt()
```

```{r}
#| tbl-cap: "Standardized (XY) factor loadings by class"
pmm_102b[["parameters"]][["stdyx.standardized"]] %>%
  filter(!grepl("ON", paramHeader)) %>%
  filter(grepl("BY", paramHeader)) %>%
  select(param, est, LatentClass) %>%
  pivot_wider(names_from = LatentClass, values_from = est, names_prefix = "class_") %>%
  gt::gt()
```

The covariance between immediate and delayed word recall was allowed to vary by class.

```{r}
#| tbl-cap: "Untandardized covariances by class"
pmm_102b[["parameters"]][["unstandardized"]] %>%
  filter(!grepl("ON", paramHeader)) %>%
  filter(grepl("WITH", paramHeader)) %>%
  mutate(param = str_c(paramHeader, " ", param),
         param = str_replace(param, "\\.", " ")) %>%
  select(param, est, LatentClass) %>%
  pivot_wider(names_from = LatentClass, values_from = est, names_prefix = "class_") %>%
  gt::gt()
```

The variance of the GCP was allowed to vary by class.

```{r}
#| tbl-cap: "Unstandardized variances by class"
pmm_102b[["parameters"]][["unstandardized"]] %>%
  filter(!grepl("ON", paramHeader)) %>%
  filter(grepl("Variances", paramHeader)) %>%
  select(param, est, LatentClass) %>%
  pivot_wider(names_from = LatentClass, values_from = est, names_prefix = "class_") %>%
  gt::gt()
```

The mean of the GCP, ADL/IADL, and subjective complaints were allowed to vary by class.

```{r}
#| tbl-cap: "Unstandardized means/intercepts/thresholds by class"
pmm_102b[["parameters"]][["unstandardized"]] %>%
  filter(!grepl("ON", paramHeader)) %>%
  filter(grepl("Thresholds", paramHeader)|
           grepl("Means", paramHeader)|
           grepl("Intercepts", paramHeader)) %>%
  filter(!grepl("Categorical.Latent.Variables", LatentClass)) %>%
  select(param, est, LatentClass) %>%
  pivot_wider(names_from = LatentClass, values_from = est, names_prefix = "class_") %>%
  gt::gt()
```





```{r}

cprob_102b <- pmm_102b[["savedata"]] %>%
  as_tibble() %>%
  janitor::clean_names()

foo <- inhcap %>%
  select(vs1hcapdxeap, cogfunction2016, consensuspaneldx) %>%
  mutate(cogfunction2016_2 = car::recode(cogfunction2016, "1:2 = 1; 3 = 2"),
         consensuspaneldx_2 = car::recode(consensuspaneldx, "1:2 = 1; 3 = 2"),
         vs1hcapdxeap2 = car::recode(vs1hcapdxeap, "1:2 = 1; 3 = 2"),
         vs1hcapdxeap2_f = factor(vs1hcapdxeap2, levels = c(1, 2), 
                                  labels = c("Non-demented", "Demented"))) %>%
  bind_cols(cprob_102b) %>%
  mutate(c2 = car::recode(c, "1:2 = 1; 3 = 2"),
         c2_f = factor(c2, levels = c(1, 2), 
                                  labels = c("Non-demented", "Demented")))

labelled::var_label(foo$cogfunction2016_2) <- "2016 Cognition Category (recoded): 1 = Normal/CIND, 2 = Demented"
labelled::var_label(foo$consensuspaneldx_2) <- "Consensus panel dx (recoded): 1 = Normal/CIND, 2 = Demented"
labelled::var_label(foo$vs1hcapdxeap2_f) <- "HRS HCAP Dementia and MCI Classification (EAP version, HRS HCAP variable set 1)"
labelled::var_label(foo$c) <- "Most likely class membership"
labelled::var_label(foo$c2) <- "Most likely class membership (recoded)"
labelled::var_label(foo$c2_f) <- "Most likely class membership (recoded)"
  
goo <- foo %>% 
  slice(rep(1:n(), each = 100))

goo <- goo %>%
  group_by(id) %>%
  mutate(i = row_number())

goo <- goo %>%
  mutate(c1 = cprob1*100,
         c2 = cprob2*100 + c1,
         c_dt = case_when(i<=c1 ~ 1,
                          i<=c2 ~ 2,
                          TRUE ~ 3),
         c_dt2 = car::recode(c_dt, "1:2 = 1; 3 = 2"),
         c_dt2_f = factor(c_dt2, levels = c(1, 2), 
                                  labels = c("Non-demented", "Demented")),
         hcap16wg100 = hcap16wg/100)

labelled::var_label(goo$c_dt2) <- "Most likely class membership (recoded)"
labelled::var_label(goo$c_dt2_f) <- "Most likely class membership (recoded)"
```


### Class assignment distribution

```{r}
#| fig-cap: "Distribution of the class probibilities by the most likely class"
cprob_102b %>%
  select(id, cprob1, cprob2, cprob3) %>%
  pivot_longer(names_to = "c", values_to = "prob", cols = c(cprob1, cprob2, cprob3)) %>%
  ggplot(aes(x = prob)) +
  geom_histogram() +
  facet_grid(c ~ .) +
  hrbrthemes::theme_ipsum()
```

```{r}
#| fig-cap: "Distribution of class probabilities by Algorithmic Dx"
inhcap %>%
  select(vs1hcapdxeap) %>%
  bind_cols(cprob_102b) %>%
  select(id, cprob1, cprob2, cprob3, c, vs1hcapdxeap) %>%
  pivot_longer(names_to = "cprob", values_to = "prob", 
               cols = c(cprob1, cprob2, cprob3)) %>%
  mutate(vs1hcapdxeap_f = factor(vs1hcapdxeap, levels = 1:3, labels = c("Normal", "MCI", "Demented")),
         cprob = str_replace(cprob, "cprob", "class_")) %>%
  ggplot(aes(x = prob, vs1hcapdxeap_f)) +
  geom_boxplot() +
  facet_grid(cprob ~ .) +
  scale_y_discrete("Algorithmic Dx") +
  scale_x_continuous("Class probibility") +
  hrbrthemes::theme_ipsum()
  


```

### Class assignment comparison

#### Comparison to Algorithmic Diagnosis

```{r}
#| tbl-cap: "Model classification vs algorithmic diagnosis, no sample weights"
gtsummary::tbl_cross(foo, vs1hcapdxeap, c)
```

```{r}
#| tbl-cap: "Kappa: Model classification vs algorithmic diagnosis, no sample weights"
psych::cohen.kappa(table(foo$vs1hcapdxeap, foo$c)) %>%
  kappa_gt()

```


```{r}
#| tbl-cap: "Model classification vs algorithmic diagnosis, no sample weights"
gtsummary::tbl_cross(foo, vs1hcapdxeap2_f, c2_f)
```


```{r}
#| tbl-cap: "Kappa: Model classification vs algorithmic diagnosis, no sample weights"
psych::cohen.kappa(table(foo$vs1hcapdxeap2, foo$c2)) %>%
  kappa_gt()
```

```{r}
#| tbl-cap: "Model classification vs algorithmic diagnosis, 100 replications, no sample weights"
gtsummary::tbl_cross(goo, vs1hcapdxeap2_f, c_dt2_f)
```


```{r}
#| tbl-cap: "Kappa: Model classification vs algorithmic diagnosis, 100 replications, no sample weights"
psych::cohen.kappa(table(goo$vs1hcapdxeap2_f, goo$c_dt2_f)) %>%
  kappa_gt()
```


```{r}
#| tbl-cap: "Model classification vs algorithmic diagnosis, 100 replications, with sample weights"
weighted_tab <- xtabs(hcap16wg100 ~ vs1hcapdxeap2_f + c_dt2_f, data = goo) %>%
  round(1) 

weighted_tab %>%
  as.data.frame.matrix() %>%
  rownames_to_column("vs1hcapdxeap2_f") %>%
  gt::gt(rowname_col = "vs1hcapdxeap2_f") %>%
  gt::tab_spanner("Most likely class membership (recoded)", 2:3) %>%
  gt::tab_stubhead("HRS HCAP Dementia and MCI Classification (EAP version, HRS HCAP variable set 1)") %>%
  gt::summary_columns(
    columns = c(`Non-demented`, `Demented`),
    fns = ~ sum(.),
    new_col_names = "Total"
  ) %>%
  gt::grand_summary_rows(
    fns = list(label = "Total", id = "total", fn = "sum"),
  )
```


```{r}
#| tbl-cap: "Kappa: Model classification vs algorithmic diagnosis, 100 replications, with sample weights"
psych::cohen.kappa(weighted_tab) %>%
  kappa_gt()
```

#### Comparison to Langa-Weir

```{r}
#| tbl-cap: "Model classification vs Langa-Weir diagnosis, 100 replications, with sample weights"
weighted_tab <- xtabs(hcap16wg100 ~ cogfunction2016_2 + c_dt2_f, data = goo) %>%
  round(1) 

weighted_tab %>%
  as.data.frame.matrix() %>%
  rownames_to_column("cogfunction2016_2") %>%
  gt::gt(rowname_col = "cogfunction2016_2") %>%
  gt::tab_spanner("Most likely class membership (recoded)", 2:3) %>%
  gt::tab_stubhead("Langa-Weir classification") %>%
  gt::summary_columns(
    columns = c(`Non-demented`, `Demented`),
    fns = ~ sum(.),
    new_col_names = "Total"
  ) %>%
  gt::grand_summary_rows(
    fns = list(label = "Total", id = "total", fn = "sum"),
  )
```


```{r}
#| tbl-cap: "Kappa: Model classification vs Langa-Weir diagnosis, 100 replications, with sample weights"
psych::cohen.kappa(weighted_tab) %>%
  kappa_gt()
```

#### Comparison to Consensus panel

```{r}
#| tbl-cap: "Model classification vs Consensus panel diagnosis, 100 replications, with sample weights"
weighted_tab <- xtabs(hcap16wg100 ~ consensuspaneldx_2 + c_dt2_f, data = goo) %>%
  round(1) 

weighted_tab %>%
  as.data.frame.matrix() %>%
  rownames_to_column("consensuspaneldx_2") %>%
  gt::gt(rowname_col = "consensuspaneldx_2") %>%
  gt::tab_spanner("Most likely class membership (recoded)", 2:3) %>%
  gt::tab_stubhead("Consensus panel classification") %>%
  gt::summary_columns(
    columns = c(`Non-demented`, `Demented`),
    fns = ~ sum(.),
    new_col_names = "Total"
  ) %>%
  gt::grand_summary_rows(
    fns = list(label = "Total", id = "total", fn = "sum"),
  )
```


```{r}
#| tbl-cap: "Kappa: Model classification vs Consensus diagnosis, 100 replications, with sample weights"
psych::cohen.kappa(weighted_tab) %>%
  kappa_gt()
```


```{r}
# 
# goo_svy <- goo %>%
#   ungroup() %>%
#   srvyr::as_survey_design(
#     weights = hcap16wg100
#   )
# 
# hoo <- survey::svytable(~vs1hcapdxeap2+ c_dt2, goo_svy, round = TRUE)
# 
# psych::cohen.kappa(hoo)
```




