

## Model - Individual item indicators

### Cognition model

This model kept the neuropsych items as separate indicators (no GCP factor).  It allowed the neuropsych items, individual functional impairment indicators, subjective memory complaints, and informant scores to vary by class.

```{r}
pmm_103b <- MplusAutomation::readModels(here::here("mplus_output", "pmm_103", "pmm_hcap_103b.out"))
pmm_103b_jorm <- MplusAutomation::readModels(here::here("mplus_output", "pmm_103_jorm", "pmm_hcap_103b_jorm.out"))
PMM_100 <- readRDS(here::here("R_objects", "PMM_100.RDS"))
inhcap <- PMM_100 %>%
  filter(inHCAP==1)

inhcap_jorm <- PMM_100 %>%
  filter(inHCAP==1) %>%
  filter(!is.na(jorm))
```

#### Parameter estimates

These are the parameter estimates to show which were constrained to be equal or allowed to vary by class.

```{r}
#| tbl-cap: "Unstandardized means/intercepts/thresholds by class"
pmm_103b[["parameters"]][["unstandardized"]] %>%
  filter(!grepl("ON", paramHeader)) %>%
  filter(grepl("Thresholds", paramHeader)|
           grepl("Means", paramHeader)|
           grepl("Intercepts", paramHeader)) %>%
  filter(!grepl("Categorical.Latent.Variables", LatentClass)) %>%
  select(param, est, LatentClass) %>%
  pivot_wider(names_from = LatentClass, values_from = est, names_prefix = "class_") %>%
  gt::gt()
```

```{r}
#| tbl-cap: "Untandardized covariances by class"
pmm_103b[["parameters"]][["unstandardized"]] %>%
  filter(!grepl("ON", paramHeader)) %>%
  filter(grepl("WITH", paramHeader)) %>%
  mutate(param = str_c(paramHeader, " ", param),
         param = str_replace(param, "\\.", " ")) %>%
  select(param, est, LatentClass) %>%
  pivot_wider(names_from = LatentClass, values_from = est, names_prefix = "class_") %>%
  gt::gt()
```


### Informant model

#### Parameter estimates

These are the parameter estimates to show which were constrained to be equal or allowed to vary by class.

```{r}
#| tbl-cap: "Unstandardized means/intercepts/thresholds by class"
pmm_103b_jorm[["parameters"]][["unstandardized"]] %>%
  filter(!grepl("ON", paramHeader)) %>%
  filter(grepl("Thresholds", paramHeader)|
           grepl("Means", paramHeader)|
           grepl("Intercepts", paramHeader)) %>%
  filter(!grepl("Categorical.Latent.Variables", LatentClass)) %>%
  select(param, est, LatentClass) %>%
  pivot_wider(names_from = LatentClass, values_from = est, names_prefix = "class_") %>%
  gt::gt()
```







```{r}

normalize_mplus_savedata <- function(savedata) {
  savedata %>%
    as_tibble() %>%
    janitor::clean_names() %>%
    rename(c = tidyselect::any_of("mlcc"))
}

cprob_103b <- normalize_mplus_savedata(pmm_103b[["savedata"]]) %>%
  select(id, stratum, secu_mplus, hcap16wg,
         cprob1, cprob2, cprob3, c)

cprob_103b_jorm <- normalize_mplus_savedata(pmm_103b_jorm[["savedata"]]) %>%
  rename(cprob1_jorm = cprob1) %>%
  rename(cprob2_jorm = cprob2) %>%
  rename(cprob3_jorm = cprob3) %>%
  rename(c_jorm = c) %>%
  rename(hcap16wg_jorm = hcap16wg) %>%
  select(id, stratum, secu_mplus, hcap16wg_jorm,
         cprob1_jorm, cprob2_jorm, cprob3_jorm, c_jorm)

cprob_103b <- cprob_103b %>%
  left_join(cprob_103b_jorm, by = c("id" = "id",
                                    "stratum" = "stratum",
                                    "secu_mplus" = "secu_mplus")
  )


# foo <- inhcap %>%
foo <- PMM_100 %>%
  select(id, vs1hcapdxeap, cogfunction2016, consensuspaneldx, 
         HCAP16WGTR, HCAP16WGTR_consensus, STRATUM, SECU, inHCAP, inConsensus) %>%
  mutate(cogfunction2016_2 = car::recode(cogfunction2016, "1:2 = 1; 3 = 2"),
         consensuspaneldx_2 = car::recode(consensuspaneldx, "1:2 = 1; 3 = 2"),
         vs1hcapdxeap2 = car::recode(vs1hcapdxeap, "1:2 = 1; 3 = 2"),
         vs1hcapdxeap2_f = factor(vs1hcapdxeap2, levels = c(1, 2), 
                                  labels = c("Non-demented", "Demented")),
         id = as.numeric(id)
         ) %>%
  left_join(cprob_103b, by = "id")  %>%
  mutate(c2 = car::recode(c, "1:2 = 1; 3 = 2"),
         c2_f = factor(c2, levels = c(1, 2), 
                                  labels = c("Non-demented", "Demented")),
         c_jorm2 = car::recode(c_jorm, "1:2 = 1; 3 = 2"),
         c_jorm2_f = factor(c_jorm2, levels = c(1, 2), 
                                  labels = c("Non-demented", "Demented")),
         c_dt = case_when(!is.na(c_jorm) ~ c_jorm,
                          !is.na(c) ~ c),
         cprob1_dt = case_when(!is.na(c_jorm) ~ cprob1_jorm,
                          !is.na(c) ~ cprob1),
         cprob2_dt = case_when(!is.na(c_jorm) ~ cprob2_jorm,
                          !is.na(c) ~ cprob2),
         cprob3_dt = case_when(!is.na(c_jorm) ~ cprob3_jorm,
                          !is.na(c) ~ cprob3),
         c_dt2 = car::recode(c_dt, "1:2 = 1; 3 = 2"),
         c_dt2_f = factor(c_dt2, levels = c(1, 2), 
                                  labels = c("Non-demented", "Demented"))
         )

labelled::var_label(foo$cogfunction2016_2) <- "2016 Cognition Category (recoded): 1 = Normal/CIND, 2 = Demented"
labelled::var_label(foo$consensuspaneldx_2) <- "Consensus panel dx (recoded): 1 = Normal/CIND, 2 = Demented"
labelled::var_label(foo$vs1hcapdxeap2_f) <- "HRS HCAP Dementia and MCI Classification (EAP version, HRS HCAP variable set 1)"

labelled::var_label(foo$c) <- "Most likely class membership (cognition model)"
labelled::var_label(foo$c2) <- "Most likely class membership (2 level) (cognition model)"
labelled::var_label(foo$c2_f) <- "Most likely class membership (2 level) (cognition model)"
labelled::var_label(foo$c_jorm) <- "Most likely class membership (Jorm model)"
labelled::var_label(foo$c_jorm2) <- "Most likely class membership (2 level) (Jorm model)"
labelled::var_label(foo$c_jorm2_f) <- "Most likely class membership (2 level) (Jorm model)"

labelled::var_label(foo$c_dt) <- "Most likely class membership (combined model)"
labelled::var_label(foo$c_dt2) <- "Most likely class membership (2 level) (combined model)"
labelled::var_label(foo$c_dt2_f) <- "Most likely class membership (2 level) (combined model)"
  

goo <- foo %>% 
  slice(rep(1:n(), each = 100))

goo <- goo %>%
  group_by(id) %>%
  mutate(i = row_number()) %>%
  ungroup()

goo <- goo %>%
  mutate(cprob1_100 = cprob1*100,
         cprob2_100 = cprob2*100 + cprob1_100,
         c_100 = case_when(i<=cprob1_100 ~ 1,
                           i<=cprob2_100 ~ 2,
                           !is.na(cprob3) ~ 3),
         c_100_2 = car::recode(c_100, "1:2 = 1; 3 = 2"),
         c_100_2_f = factor(c_100_2, levels = c(1, 2), 
                            labels = c("Non-demented", "Demented")),
         cprob1_jorm_100 = cprob1_jorm*100,
         cprob2_jorm_100 = cprob2_jorm*100 + cprob1_jorm_100,
         c_jorm_100 = case_when(i<=cprob1_jorm_100 ~ 1,
                                i<=cprob2_jorm_100 ~ 2,
                                !is.na(cprob3_jorm) ~ 3),
         c_jorm_100_2 = car::recode(c_jorm_100, "1:2 = 1; 3 = 2"),
         c_jorm_100_2_f = factor(c_jorm_100_2, levels = c(1, 2), 
                                  labels = c("Non-demented", "Demented")),
         cprob1_dt_100 = cprob1_dt*100,
         cprob2_dt_100 = cprob2_dt*100 + cprob1_dt_100,
         c_dt_100 = case_when(i<=cprob1_dt_100 ~ 1,
                              i<=cprob2_dt_100 ~ 2,
                              !is.na(cprob3_dt) ~ 3),
         c_dt_100_2 = car::recode(c_dt_100, "1:2 = 1; 3 = 2"),
         c_dt_100_2_f = factor(c_dt_100_2, levels = c(1, 2), 
                                  labels = c("Non-demented", "Demented")),
         hcap16wg100 = case_when(!is.na(HCAP16WGTR) ~ HCAP16WGTR /100,
                                 is.na(HCAP16WGTR) ~ 0),
         hcap16wg_consensus100 = case_when(!is.na(HCAP16WGTR_consensus) ~ HCAP16WGTR_consensus /100,
                                 is.na(HCAP16WGTR_consensus) ~ 0)
  )

labelled::var_label(goo$c_100) <- "Most likely class membership (cognition model)"
labelled::var_label(goo$c_100_2) <- "Most likely class membership (2 level) (cognition model)"
labelled::var_label(goo$c_100_2_f) <- "Most likely class membership (2 level) (cognition model)"
labelled::var_label(goo$c_jorm_100) <- "Most likely class membership (Jorm model)"
labelled::var_label(goo$c_jorm_100_2) <- "Most likely class membership (2 level) (Jorm model)"
labelled::var_label(goo$c_jorm_100_2_f) <- "Most likely class membership (2 level) (Jorm model)"

labelled::var_label(goo$c_dt_100) <- "Most likely class membership (combined model)"
labelled::var_label(goo$c_dt_100_2) <- "Most likely class membership (2 level) (combined model)"
labelled::var_label(goo$c_dt_100_2_f) <- "Most likely class membership (2 level) (combined model)"
```

```{r}
hrs_design100 <- survey::svydesign(
  ids = ~id,
  strata = ~STRATUM,
  weights = ~hcap16wg100,
  data = goo,
  nest = TRUE
)

# Restrict analysis to a subpopulation (e.g., age 65+)
hcap_design100 <- subset(hrs_design100, inHCAP==1)
consensus_design100 <- survey::svydesign(
  ids = ~id,
  strata = ~STRATUM,
  weights = ~hcap16wg_consensus100,
  data = goo,
  nest = TRUE
)
consensus_design100 <- subset(consensus_design100, inConsensus==1)

```

### Class probability distribution

```{r}
#| fig-cap: "Distribution of the class probibilities from the cognition model by the most likely class"
cprob_103b %>%
  select(id, cprob1, cprob2, cprob3) %>%
  pivot_longer(names_to = "c", values_to = "prob", cols = c(cprob1, cprob2, cprob3)) %>%
  ggplot(aes(x = prob)) +
  geom_histogram() +
  facet_grid(c ~ .) +
  hrbrthemes::theme_ipsum()
```


```{r}
#| fig-cap: "Distribution of class probabilities from the cognition model by Algorithmic Dx"

foo %>%
  filter(!is.na(c)) %>%
  select(id, cprob1, cprob2, cprob3, c, vs1hcapdxeap) %>%
  pivot_longer(names_to = "cprob", values_to = "prob", 
               cols = c(cprob1, cprob2, cprob3)) %>%
  mutate(vs1hcapdxeap_f = factor(vs1hcapdxeap, levels = 1:3, labels = c("Normal", "MCI", "Demented")),
         cprob = str_replace(cprob, "cprob", "class_")) %>%
  ggplot(aes(x = prob, vs1hcapdxeap_f)) +
  geom_boxplot() +
  facet_grid(cprob ~ .) +
  scale_y_discrete("Algorithmic Dx") +
  scale_x_continuous("Class probibility") +
  hrbrthemes::theme_ipsum()
  


```

```{r}
#| fig-cap: "Distribution of the class probibilities from the Jorm model by the most likely class"
cprob_103b %>%
  select(id, cprob1_jorm, cprob2_jorm, cprob3_jorm) %>%
  pivot_longer(names_to = "c", values_to = "prob", cols = c(cprob1_jorm, cprob2_jorm, cprob3_jorm)) %>%
  ggplot(aes(x = prob)) +
  geom_histogram() +
  facet_grid(c ~ .) +
  hrbrthemes::theme_ipsum()
```


```{r}
#| fig-cap: "Distribution of class probabilities from the Jorm model by Algorithmic Dx"

foo %>%
  filter(!is.na(c_jorm)) %>%
  select(id, cprob1_jorm, cprob2_jorm, cprob3_jorm, c_jorm, vs1hcapdxeap) %>%
  pivot_longer(names_to = "cprob", values_to = "prob", 
               cols = c(cprob1_jorm, cprob2_jorm, cprob3_jorm)) %>%
  mutate(vs1hcapdxeap_f = factor(vs1hcapdxeap, levels = 1:3, labels = c("Normal", "MCI", "Demented")),
         cprob = str_replace(cprob, "cprob", "class_")) %>%
  ggplot(aes(x = prob, vs1hcapdxeap_f)) +
  geom_boxplot() +
  facet_grid(cprob ~ .) +
  scale_y_discrete("Algorithmic Dx") +
  scale_x_continuous("Class probibility") +
  hrbrthemes::theme_ipsum()
  


```

### Class assignment comparison

#### Comparison to Algorithmic Diagnosis

**Cognition model**

```{r}
#| tbl-cap: "Most likely class membership (cognition model) vs algorithmic diagnosis, no sample weights"

foo %>%
  filter(!is.na(c)) %>%
  gtsummary::tbl_cross(vs1hcapdxeap, c)
```


```{r}
#| tbl-cap: "Kappa: Most likely class membership (cognition model) vs algorithmic diagnosis, no sample weights"

foo %>%
  filter(!is.na(c)) %>%
  select(vs1hcapdxeap, c) %>%
  table() %>%
  psych::cohen.kappa() %>%
  kappa_gt()

```


```{r}
#| tbl-cap: "Most likely class membership (cognition model) vs algorithmic diagnosis, no sample weights"

foo %>%
  filter(!is.na(c)) %>%
  gtsummary::tbl_cross(vs1hcapdxeap2_f, c2_f)
```


```{r}
#| tbl-cap: "Kappa: Most likely class membership (cognition model) vs algorithmic diagnosis, no sample weights"

foo %>%
  filter(!is.na(c)) %>%
  select(vs1hcapdxeap2, c2) %>%
  table() %>%
  psych::cohen.kappa() %>%
  kappa_gt()

```

**Jorm model**

```{r}
#| tbl-cap: "Most likely class membership (Jorm model) vs algorithmic diagnosis, no sample weights"

foo %>%
  filter(!is.na(c_jorm)) %>%
  gtsummary::tbl_cross(vs1hcapdxeap, c_jorm)
```


```{r}
#| tbl-cap: "Kappa: Most likely class membership (Jorm model) vs algorithmic diagnosis, no sample weights"

foo %>%
  filter(!is.na(c_jorm)) %>%
  select(vs1hcapdxeap, c_jorm) %>%
  table() %>%
  psych::cohen.kappa() %>%
  kappa_gt()
```


```{r}
#| tbl-cap: "Most likely class membership (Jorm model) vs algorithmic diagnosis, no sample weights"

foo %>%
  filter(!is.na(c_jorm)) %>%
  gtsummary::tbl_cross(vs1hcapdxeap2_f, c_jorm2_f)
```


```{r}
#| tbl-cap: "Kappa: Most likely class membership (Jorm model) vs algorithmic diagnosis, no sample weights"

foo %>%
  filter(!is.na(c_jorm)) %>%
  select(vs1hcapdxeap2, c_jorm2) %>%
  table() %>%
  psych::cohen.kappa() %>%
  kappa_gt()

```

```{r}
#| tbl-cap: "Comparison of class membership between the cognition model and Jorm model"
foo %>%
  filter(!is.na(c)) %>%
  gtsummary::tbl_cross(c, c_jorm)
```

**Combined model**

```{r}
#| tbl-cap: "Most likely class membership (combined model) vs algorithmic diagnosis, no sample weights"
foo %>%
  filter(!is.na(c)) %>%
  gtsummary::tbl_cross(vs1hcapdxeap, c_dt)
```


```{r}
#| tbl-cap: "Kappa: Most likely class membership (combined model) vs algorithmic diagnosis, no sample weights"

foo %>%
  filter(!is.na(c)) %>%
  select(vs1hcapdxeap, c_dt) %>%
  table() %>%
  psych::cohen.kappa() %>%
  kappa_gt()
```


```{r}
#| tbl-cap: "Most likely class membership (combined model) vs algorithmic diagnosis, no sample weights"
foo %>%
  filter(!is.na(c)) %>%
  gtsummary::tbl_cross(vs1hcapdxeap2_f, c_dt2_f)
```


```{r}
#| tbl-cap: "Kappa: Most likely class membership (combined model) vs algorithmic diagnosis, no sample weights"

foo %>%
  filter(!is.na(c)) %>%
  select(vs1hcapdxeap2, c_dt2) %>%
  table() %>%
  psych::cohen.kappa() %>%
  kappa_gt()

```

```{r}
#| tbl-cap: "Most likely class membership (combined model) vs algorithmic diagnosis, 100 replications, no sample weights"

goo %>%
  filter(!is.na(c_dt)) %>%
  gtsummary::tbl_cross(vs1hcapdxeap2_f, c_dt_100_2_f)
```


```{r}
#| tbl-cap: "Kappa: Most likely class membership (combined model) vs algorithmic diagnosis, 100 replications, no sample weights"

goo %>%
  filter(!is.na(c_dt)) %>%
  select(vs1hcapdxeap2, c_dt_100_2) %>%
  table() %>%
  psych::cohen.kappa() %>%
  kappa_gt()

```



```{r}
#| tbl-cap: "Most likely class membership (cognition model) vs algorithmic diagnosis, 100 replications, with sample weights"

weighted_tab <- svytable(~ vs1hcapdxeap2_f + c_dt2_f, hcap_design100)


weighted_tab %>%
  as.data.frame.matrix() %>%
  rownames_to_column("vs1hcapdxeap2_f") %>%
  gt::gt(rowname_col = "vs1hcapdxeap2_f") %>%
  gt::fmt_number(decimals = 0) %>%
  gt::tab_spanner("Most likely class membership (combined model)", 2:3) %>%
  gt::tab_stubhead("HRS HCAP Dementia and MCI Classification (EAP version, HRS HCAP variable set 1)") %>%
  gt::summary_columns(
    columns = c(`Non-demented`, `Demented`),
    fns = ~ sum(.),
    new_col_names = "Total",
    fmt = ~gt::fmt_number(., decimals = 0),
  ) %>%
  gt::grand_summary_rows(
    fns = list(label = "Total", id = "total", fn = "sum"),
    fmt = ~gt::fmt_number(., decimals = 0)
  )
```


```{r}
#| tbl-cap: "Kappa: Most likely class membership (cognition model) vs algorithmic diagnosis, 100 replications, with sample weights"
psych::cohen.kappa(weighted_tab) %>%
  kappa_gt()
```


#### Comparison to Langa-Weir

```{r}
#| tbl-cap: "Model classification vs Langa-Weir diagnosis, 100 replications, with sample weights"
weighted_tab <- svytable(~ cogfunction2016_2 + c_dt2_f, hcap_design100)


weighted_tab %>%
  as.data.frame.matrix() %>%
  rownames_to_column("cogfunction2016_2") %>%
  gt::gt(rowname_col = "cogfunction2016_2") %>%
  gt::fmt_number(decimals = 0) %>%
  gt::tab_spanner("Most likely class membership (combined model)", 2:3) %>%
  gt::tab_stubhead("Langa-Weir classification") %>%
  gt::summary_columns(
    columns = c(`Non-demented`, `Demented`),
    fns = ~ sum(.),
    new_col_names = "Total",
    fmt = ~gt::fmt_number(., decimals = 0),
  ) %>%
  gt::grand_summary_rows(
    fns = list(label = "Total", id = "total", fn = "sum"),
    fmt = ~gt::fmt_number(., decimals = 0)
  )
```


```{r}
#| tbl-cap: "Kappa: Model classification vs Langa-Weir diagnosis, 100 replications, with sample weights"
psych::cohen.kappa(weighted_tab) %>%
  kappa_gt()
```

#### Comparison to Consensus panel

```{r}
#| tbl-cap: "Model classification vs Consensus panel diagnosis, 100 replications, with sample weights"

weighted_tab <- svytable(~ consensuspaneldx_2 + c_dt2_f, consensus_design100) %>%
  round(0) 


weighted_tab %>%
  as.data.frame.matrix() %>%
  rownames_to_column("consensuspaneldx_2") %>%
  gt::gt(rowname_col = "consensuspaneldx_2") %>%
  gt::fmt_number(decimals = 0) %>%
  gt::tab_spanner("Most likely class membership (combined model)", 2:3) %>%
  gt::tab_stubhead("Consensus panel classification") %>%
  gt::summary_columns(
    columns = c(`Non-demented`, `Demented`),
    fns = ~ sum(.),
    new_col_names = "Total",
    fmt = ~gt::fmt_number(., decimals = 0),
  ) %>%
  gt::grand_summary_rows(
    fns = list(label = "Total", id = "total", fn = "sum"),
    fmt = ~gt::fmt_number(., decimals = 0)
  )
```


```{r}
#| tbl-cap: "Kappa: Model classification vs Consensus diagnosis, 100 replications, with sample weights"
psych::cohen.kappa(weighted_tab) %>%
  kappa_gt()
```




