```{r}
#| include: false

source(here::here("R", "PMM_112_Margins.R"))

style_margins_slide_plot <- function(plot_obj) {
  plot_obj +
    ggplot2::theme_minimal(base_size = 20) +
    ggplot2::theme(
      panel.grid.minor = ggplot2::element_blank(),
      strip.text = ggplot2::element_text(face = "bold", size = 18)
    )
}

style_margins_slide_table <- function(gt_tbl) {
  gt_tbl |>
    gt::opt_stylize(6, color = "gray") |>
    gt::tab_options(
      table.width = gt::pct(80),
      table.font.size = gt::px(24),
      column_labels.font.size = gt::px(22),
      stub.font.size = gt::px(22),
      source_notes.font.size = gt::px(16),
      data_row.padding = gt::px(4)
    )
}
```

## Combined-Model Class Probabilities

```{r}
#| fig-width: 10
#| fig-height: 5
style_margins_slide_plot(PMM_112$figure)
```

---

## Combined-Model Class Membership 

<div style="font-size: 0.8em; line-height: 1.1;">
According to most likely class membership. This is the value that would be used to generate a "predicted" class label for each participant in the HRS/HCAP sample.
</div>

```{r}
style_margins_slide_table(PMM_112$table)
```


<div style="font-size: 0.66em; line-height: 0.9;">
By way of comparison, Manly et al. (2022) report that in the HCAP validation subsample, 68% of participants were classified as normal, 22% as MCI, and 10% as dementia according to the HRS/HCAP algorithm (weighted %, class assignments based on Bayesian Plausible Values for cognition scores). Langa-Weir classification of the HRS/HCAP sample classifies 75% as normal, 17% as MCI, and 8% as dementia (see PMM_Analysis_Report).
</div>
