```{r}
#| include: false

PMM_100 <- readRDS(here::here("R_objects", "PMM_100.RDS"))
algo_core <- readRDS(here::here("R_objects", "030_hcap.rds"))

class_labels <- c("Normal", "MCI", "Dementia")
binary_labels <- c("Non-demented", "Demented")
agreement_methods <- c(
  "Consensus diagnosis",
  "Algorithmic diagnosis",
  "Langa-Weir",
  "HCAP Algo in Core",
  "PMM"
)

weighted_kappa_from_tab <- function(tab) {
  obs <- unclass(tab)
  expected <- outer(rowSums(obs), colSums(obs)) / sum(obs)

  agreement_weights <- outer(
    1:nrow(obs),
    1:ncol(obs),
    function(i, j) 1 - ((i - j)^2 / (nrow(obs) - 1)^2)
  )

  po <- sum(agreement_weights * obs) / sum(obs)
  pe <- sum(agreement_weights * expected) / sum(obs)
  kappa_w <- (po - pe) / (1 - pe)

  tibble::tibble(
    observed_weighted_agreement = po,
    expected_weighted_agreement = pe,
    quadratic_weighted_kappa = kappa_w
  )
}

binary_kappa_from_tab <- function(tab) {
  obs <- unclass(tab)
  expected <- outer(rowSums(obs), colSums(obs)) / sum(obs)

  po <- sum(diag(obs)) / sum(obs)
  pe <- sum(diag(expected)) / sum(obs)
  kappa <- (po - pe) / (1 - pe)

  tibble::tibble(
    observed_agreement = po,
    expected_agreement = pe,
    kappa = kappa
  )
}

style_slide_gt <- function(gt_tbl) {
  gt_tbl |>
    gt::cols_width(
      gt::everything() ~ gt::px(120)
    ) |>
    gt::opt_stylize(6, color = "gray") |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(24),
      column_labels.font.size = gt::px(21),
      stub.font.size = gt::px(21),
      source_notes.font.size = gt::px(19),
      data_row.padding = gt::px(3)
    )
}

format_weighted_tab <- function(tab, row_name, stubhead, spanner) {
  tab_df <- tab |>
    as.data.frame.matrix() |>
    tibble::rownames_to_column(row_name) |>
    dplyr::mutate(Total = Normal + MCI + Dementia)

  total_row <- tab_df |>
    dplyr::summarise(
      dplyr::across(c(Normal, MCI, Dementia, Total), sum),
      .groups = "drop"
    ) |>
    dplyr::mutate(!!row_name := "Total") |>
    dplyr::select(dplyr::all_of(c(row_name, "Normal", "MCI", "Dementia", "Total")))

  dplyr::bind_rows(tab_df, total_row) |>
    gt::gt(rowname_col = row_name) |>
    gt::tab_spanner(spanner, 2:4) |>
    gt::tab_stubhead(stubhead)
}

format_binary_weighted_tab <- function(tab, row_name, stubhead, spanner) {
  tab_df <- tab |>
    as.data.frame.matrix() |>
    tibble::rownames_to_column(row_name) |>
    dplyr::mutate(Total = `Non-demented` + Demented)

  total_row <- tab_df |>
    dplyr::summarise(
      dplyr::across(c(`Non-demented`, Demented, Total), sum),
      .groups = "drop"
    ) |>
    dplyr::mutate(!!row_name := "Total") |>
    dplyr::select(dplyr::all_of(c(row_name, "Non-demented", "Demented", "Total")))

  dplyr::bind_rows(tab_df, total_row) |>
    gt::gt(rowname_col = row_name) |>
    gt::tab_spanner(spanner, 2:3) |>
    gt::tab_stubhead(stubhead)
}

add_kappa_note <- function(gt_tbl, label, value) {
  gt_tbl |>
    gt::tab_source_note(
      gt::md(stringr::str_glue("{label}: **{scales::number(value, accuracy = 0.001)}**"))
    )
}

consensus_sample <- PMM_100 |>
  dplyr::filter(
    inHCAP == 1,
    !is.na(consensuspaneldx),
    !is.na(cogfunction2016),
    !is.na(vs1hcapdxeap),
    !is.na(samplingP)
  ) |>
  dplyr::transmute(
    HHID,
    PN,
    id = as.integer(id),
    consensus = haven::zap_labels(consensuspaneldx),
    algorithmic = haven::zap_labels(vs1hcapdxeap),
    langa_weir = haven::zap_labels(cogfunction2016),
    consensus_wt_raw = HCAP16WGTR * samplingP^(-1)
  ) |>
  dplyr::left_join(
    algo_core |>
      dplyr::transmute(
        HHID,
        PN,
        core_algo_raw = haven::zap_labels(dx_v2)
      ),
    by = c("HHID", "PN")
  ) |>
  dplyr::mutate(
    consensus_wt = 100 * consensus_wt_raw / sum(consensus_wt_raw, na.rm = TRUE),
    consensus = factor(consensus, levels = 1:3, labels = class_labels, ordered = TRUE),
    algorithmic = factor(algorithmic, levels = 1:3, labels = class_labels, ordered = TRUE),
    langa_weir = factor(langa_weir, levels = 1:3, labels = class_labels, ordered = TRUE),
    core_algo = factor(core_algo_raw + 1, levels = 1:3, labels = class_labels, ordered = TRUE)
  )

consensus_sample_binary <- consensus_sample |>
  dplyr::mutate(
    consensus_binary = factor(
      dplyr::if_else(as.integer(consensus) == 3L, 2L, 1L),
      levels = 1:2,
      labels = binary_labels,
      ordered = TRUE
    ),
    algorithmic_binary = factor(
      dplyr::if_else(as.integer(algorithmic) == 3L, 2L, 1L),
      levels = 1:2,
      labels = binary_labels,
      ordered = TRUE
    ),
    langa_weir_binary = factor(
      dplyr::if_else(as.integer(langa_weir) == 3L, 2L, 1L),
      levels = 1:2,
      labels = binary_labels,
      ordered = TRUE
    ),
    core_algo_binary = factor(
      dplyr::if_else(as.integer(core_algo) == 3L, 2L, 1L),
      levels = 1:2,
      labels = binary_labels,
      ordered = TRUE
    )
  )

pmm_103_cprob <- MplusAutomation::readModels(
  here::here("mplus_output", "pmm_103", "pmm_hcap_103b.out")
)[["savedata"]] |>
  tibble::as_tibble() |>
  janitor::clean_names() |>
  dplyr::select(id, cprob1, cprob2, cprob3)

consensus_sample_pmm <- consensus_sample |>
  dplyr::inner_join(pmm_103_cprob, by = "id")

consensus_sample_pmm_rep <- consensus_sample_pmm |>
  dplyr::slice(rep(1:dplyr::n(), each = 100)) |>
  dplyr::group_by(id) |>
  dplyr::mutate(
    i = dplyr::row_number(),
    p1 = cprob1 * 100,
    p2 = cprob2 * 100 + p1,
    pmm = dplyr::case_when(
      i <= p1 ~ 1L,
      i <= p2 ~ 2L,
      TRUE ~ 3L
    ),
    consensus_wt100 = consensus_wt / 100
  ) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    pmm = factor(pmm, levels = 1:3, labels = class_labels, ordered = TRUE)
  )

consensus_sample_pmm_binary <- consensus_sample_pmm_rep |>
  dplyr::mutate(
    consensus_binary = factor(
      dplyr::if_else(as.integer(consensus) == 3L, 2L, 1L),
      levels = 1:2,
      labels = binary_labels,
      ordered = TRUE
    ),
    algorithmic_binary = factor(
      dplyr::if_else(as.integer(algorithmic) == 3L, 2L, 1L),
      levels = 1:2,
      labels = binary_labels,
      ordered = TRUE
    ),
    langa_weir_binary = factor(
      dplyr::if_else(as.integer(langa_weir) == 3L, 2L, 1L),
      levels = 1:2,
      labels = binary_labels,
      ordered = TRUE
    ),
    core_algo_binary = factor(
      dplyr::if_else(as.integer(core_algo) == 3L, 2L, 1L),
      levels = 1:2,
      labels = binary_labels,
      ordered = TRUE
    ),
    pmm_binary = factor(
      dplyr::if_else(as.integer(pmm) == 3L, 2L, 1L),
      levels = 1:2,
      labels = binary_labels,
      ordered = TRUE
    )
  )

tab_lw_consensus <- xtabs(consensus_wt ~ langa_weir + consensus, data = consensus_sample) |> round(1)
tab_algorithmic_consensus <- xtabs(consensus_wt ~ algorithmic + consensus, data = consensus_sample) |> round(1)
tab_algorithmic_lw <- xtabs(consensus_wt ~ algorithmic + langa_weir, data = consensus_sample) |> round(1)
tab_core_algo_consensus <- xtabs(consensus_wt ~ core_algo + consensus, data = consensus_sample) |> round(1)
tab_core_algo_algorithmic <- xtabs(consensus_wt ~ core_algo + algorithmic, data = consensus_sample) |> round(1)
tab_core_algo_lw <- xtabs(consensus_wt ~ core_algo + langa_weir, data = consensus_sample) |> round(1)
tab_consensus_pmm <- xtabs(consensus_wt100 ~ consensus + pmm, data = consensus_sample_pmm_rep) |> round(1)
tab_algorithmic_pmm <- xtabs(consensus_wt100 ~ algorithmic + pmm, data = consensus_sample_pmm_rep) |> round(1)
tab_lw_pmm <- xtabs(consensus_wt100 ~ langa_weir + pmm, data = consensus_sample_pmm_rep) |> round(1)
tab_core_algo_pmm <- xtabs(consensus_wt100 ~ core_algo + pmm, data = consensus_sample_pmm_rep) |> round(1)

kappa_lw_consensus <- weighted_kappa_from_tab(tab_lw_consensus)
kappa_algorithmic_consensus <- weighted_kappa_from_tab(tab_algorithmic_consensus)
kappa_algorithmic_lw <- weighted_kappa_from_tab(tab_algorithmic_lw)
kappa_core_algo_consensus <- weighted_kappa_from_tab(tab_core_algo_consensus)
kappa_core_algo_algorithmic <- weighted_kappa_from_tab(tab_core_algo_algorithmic)
kappa_core_algo_lw <- weighted_kappa_from_tab(tab_core_algo_lw)
kappa_consensus_pmm <- weighted_kappa_from_tab(tab_consensus_pmm)
kappa_algorithmic_pmm <- weighted_kappa_from_tab(tab_algorithmic_pmm)
kappa_lw_pmm <- weighted_kappa_from_tab(tab_lw_pmm)
kappa_core_algo_pmm <- weighted_kappa_from_tab(tab_core_algo_pmm)

agreement_matrix <- matrix(
  NA_real_,
  nrow = length(agreement_methods),
  ncol = length(agreement_methods),
  dimnames = list(agreement_methods, agreement_methods)
)

diag(agreement_matrix) <- 1
agreement_matrix["Consensus diagnosis", "Algorithmic diagnosis"] <- kappa_algorithmic_consensus$quadratic_weighted_kappa
agreement_matrix["Algorithmic diagnosis", "Consensus diagnosis"] <- kappa_algorithmic_consensus$quadratic_weighted_kappa
agreement_matrix["Consensus diagnosis", "Langa-Weir"] <- kappa_lw_consensus$quadratic_weighted_kappa
agreement_matrix["Langa-Weir", "Consensus diagnosis"] <- kappa_lw_consensus$quadratic_weighted_kappa
agreement_matrix["Consensus diagnosis", "HCAP Algo in Core"] <- kappa_core_algo_consensus$quadratic_weighted_kappa
agreement_matrix["HCAP Algo in Core", "Consensus diagnosis"] <- kappa_core_algo_consensus$quadratic_weighted_kappa
agreement_matrix["Consensus diagnosis", "PMM"] <- kappa_consensus_pmm$quadratic_weighted_kappa
agreement_matrix["PMM", "Consensus diagnosis"] <- kappa_consensus_pmm$quadratic_weighted_kappa
agreement_matrix["Algorithmic diagnosis", "Langa-Weir"] <- kappa_algorithmic_lw$quadratic_weighted_kappa
agreement_matrix["Langa-Weir", "Algorithmic diagnosis"] <- kappa_algorithmic_lw$quadratic_weighted_kappa
agreement_matrix["Algorithmic diagnosis", "HCAP Algo in Core"] <- kappa_core_algo_algorithmic$quadratic_weighted_kappa
agreement_matrix["HCAP Algo in Core", "Algorithmic diagnosis"] <- kappa_core_algo_algorithmic$quadratic_weighted_kappa
agreement_matrix["Algorithmic diagnosis", "PMM"] <- kappa_algorithmic_pmm$quadratic_weighted_kappa
agreement_matrix["PMM", "Algorithmic diagnosis"] <- kappa_algorithmic_pmm$quadratic_weighted_kappa
agreement_matrix["Langa-Weir", "HCAP Algo in Core"] <- kappa_core_algo_lw$quadratic_weighted_kappa
agreement_matrix["HCAP Algo in Core", "Langa-Weir"] <- kappa_core_algo_lw$quadratic_weighted_kappa
agreement_matrix["Langa-Weir", "PMM"] <- kappa_lw_pmm$quadratic_weighted_kappa
agreement_matrix["PMM", "Langa-Weir"] <- kappa_lw_pmm$quadratic_weighted_kappa
agreement_matrix["HCAP Algo in Core", "PMM"] <- kappa_core_algo_pmm$quadratic_weighted_kappa
agreement_matrix["PMM", "HCAP Algo in Core"] <- kappa_core_algo_pmm$quadratic_weighted_kappa

tab_lw_consensus_binary <- xtabs(consensus_wt ~ langa_weir_binary + consensus_binary, data = consensus_sample_binary) |> round(1)
tab_algorithmic_consensus_binary <- xtabs(consensus_wt ~ algorithmic_binary + consensus_binary, data = consensus_sample_binary) |> round(1)
tab_algorithmic_lw_binary <- xtabs(consensus_wt ~ algorithmic_binary + langa_weir_binary, data = consensus_sample_binary) |> round(1)
tab_core_algo_consensus_binary <- xtabs(consensus_wt ~ core_algo_binary + consensus_binary, data = consensus_sample_binary) |> round(1)
tab_core_algo_algorithmic_binary <- xtabs(consensus_wt ~ core_algo_binary + algorithmic_binary, data = consensus_sample_binary) |> round(1)
tab_core_algo_lw_binary <- xtabs(consensus_wt ~ core_algo_binary + langa_weir_binary, data = consensus_sample_binary) |> round(1)
tab_consensus_pmm_binary <- xtabs(consensus_wt100 ~ consensus_binary + pmm_binary, data = consensus_sample_pmm_binary) |> round(1)
tab_algorithmic_pmm_binary <- xtabs(consensus_wt100 ~ algorithmic_binary + pmm_binary, data = consensus_sample_pmm_binary) |> round(1)
tab_lw_pmm_binary <- xtabs(consensus_wt100 ~ langa_weir_binary + pmm_binary, data = consensus_sample_pmm_binary) |> round(1)
tab_core_algo_pmm_binary <- xtabs(consensus_wt100 ~ core_algo_binary + pmm_binary, data = consensus_sample_pmm_binary) |> round(1)

kappa_lw_consensus_binary <- binary_kappa_from_tab(tab_lw_consensus_binary)
kappa_algorithmic_consensus_binary <- binary_kappa_from_tab(tab_algorithmic_consensus_binary)
kappa_algorithmic_lw_binary <- binary_kappa_from_tab(tab_algorithmic_lw_binary)
kappa_core_algo_consensus_binary <- binary_kappa_from_tab(tab_core_algo_consensus_binary)
kappa_core_algo_algorithmic_binary <- binary_kappa_from_tab(tab_core_algo_algorithmic_binary)
kappa_core_algo_lw_binary <- binary_kappa_from_tab(tab_core_algo_lw_binary)
kappa_consensus_pmm_binary <- binary_kappa_from_tab(tab_consensus_pmm_binary)
kappa_algorithmic_pmm_binary <- binary_kappa_from_tab(tab_algorithmic_pmm_binary)
kappa_lw_pmm_binary <- binary_kappa_from_tab(tab_lw_pmm_binary)
kappa_core_algo_pmm_binary <- binary_kappa_from_tab(tab_core_algo_pmm_binary)

agreement_matrix_binary <- matrix(
  NA_real_,
  nrow = length(agreement_methods),
  ncol = length(agreement_methods),
  dimnames = list(agreement_methods, agreement_methods)
)

diag(agreement_matrix_binary) <- 1
agreement_matrix_binary["Consensus diagnosis", "Algorithmic diagnosis"] <- kappa_algorithmic_consensus_binary$kappa
agreement_matrix_binary["Algorithmic diagnosis", "Consensus diagnosis"] <- kappa_algorithmic_consensus_binary$kappa
agreement_matrix_binary["Consensus diagnosis", "Langa-Weir"] <- kappa_lw_consensus_binary$kappa
agreement_matrix_binary["Langa-Weir", "Consensus diagnosis"] <- kappa_lw_consensus_binary$kappa
agreement_matrix_binary["Consensus diagnosis", "HCAP Algo in Core"] <- kappa_core_algo_consensus_binary$kappa
agreement_matrix_binary["HCAP Algo in Core", "Consensus diagnosis"] <- kappa_core_algo_consensus_binary$kappa
agreement_matrix_binary["Consensus diagnosis", "PMM"] <- kappa_consensus_pmm_binary$kappa
agreement_matrix_binary["PMM", "Consensus diagnosis"] <- kappa_consensus_pmm_binary$kappa
agreement_matrix_binary["Algorithmic diagnosis", "Langa-Weir"] <- kappa_algorithmic_lw_binary$kappa
agreement_matrix_binary["Langa-Weir", "Algorithmic diagnosis"] <- kappa_algorithmic_lw_binary$kappa
agreement_matrix_binary["Algorithmic diagnosis", "HCAP Algo in Core"] <- kappa_core_algo_algorithmic_binary$kappa
agreement_matrix_binary["HCAP Algo in Core", "Algorithmic diagnosis"] <- kappa_core_algo_algorithmic_binary$kappa
agreement_matrix_binary["Algorithmic diagnosis", "PMM"] <- kappa_algorithmic_pmm_binary$kappa
agreement_matrix_binary["PMM", "Algorithmic diagnosis"] <- kappa_algorithmic_pmm_binary$kappa
agreement_matrix_binary["Langa-Weir", "HCAP Algo in Core"] <- kappa_core_algo_lw_binary$kappa
agreement_matrix_binary["HCAP Algo in Core", "Langa-Weir"] <- kappa_core_algo_lw_binary$kappa
agreement_matrix_binary["Langa-Weir", "PMM"] <- kappa_lw_pmm_binary$kappa
agreement_matrix_binary["PMM", "Langa-Weir"] <- kappa_lw_pmm_binary$kappa
agreement_matrix_binary["HCAP Algo in Core", "PMM"] <- kappa_core_algo_pmm_binary$kappa
agreement_matrix_binary["PMM", "HCAP Algo in Core"] <- kappa_core_algo_pmm_binary$kappa
```

## Three-class: Pairwise Kappa Matrix

```{r}
agreement_matrix |>
  as.data.frame() |>
  tibble::rownames_to_column("method") |>
  gt::gt(rowname_col = "method") |>
  gt::fmt_number(columns = agreement_methods, decimals = 3) |>
  gt::tab_stubhead("Consensus sample") |>
  style_slide_gt()
```

## Binary: Pairwise Kappa Matrix

```{r}
agreement_matrix_binary |>
  as.data.frame() |>
  tibble::rownames_to_column("method") |>
  gt::gt(rowname_col = "method") |>
  gt::fmt_number(columns = agreement_methods, decimals = 3) |>
  gt::tab_stubhead("Consensus sample") |>
  style_slide_gt()
```

## Three-class: Langa-Weir vs Consensus

```{r}
format_weighted_tab(
  tab_lw_consensus,
  row_name = "langa_weir",
  stubhead = "Langa-Weir classification",
  spanner = "Consensus panel diagnosis"
) |>
  add_kappa_note("Quadratic weighted kappa", kappa_lw_consensus$quadratic_weighted_kappa) |>
  style_slide_gt()
```

## Three-class: Algorithmic vs Consensus

```{r}
format_weighted_tab(
  tab_algorithmic_consensus,
  row_name = "algorithmic",
  stubhead = "Algorithmic diagnosis",
  spanner = "Consensus panel diagnosis"
) |>
  add_kappa_note("Quadratic weighted kappa", kappa_algorithmic_consensus$quadratic_weighted_kappa) |>
  style_slide_gt()
```

## Three-class: Algorithmic vs Langa-Weir

```{r}
format_weighted_tab(
  tab_algorithmic_lw,
  row_name = "algorithmic",
  stubhead = "Algorithmic diagnosis",
  spanner = "Langa-Weir diagnosis"
) |>
  add_kappa_note("Quadratic weighted kappa", kappa_algorithmic_lw$quadratic_weighted_kappa) |>
  style_slide_gt()
```

## Three-class: HCAP Algo in Core vs Consensus

```{r}
format_weighted_tab(
  tab_core_algo_consensus,
  row_name = "core_algo",
  stubhead = "HCAP Algo in Core",
  spanner = "Consensus panel diagnosis"
) |>
  add_kappa_note("Quadratic weighted kappa", kappa_core_algo_consensus$quadratic_weighted_kappa) |>
  style_slide_gt()
```

## Three-class: HCAP Algo in Core vs Algorithmic

```{r}
format_weighted_tab(
  tab_core_algo_algorithmic,
  row_name = "core_algo",
  stubhead = "HCAP Algo in Core",
  spanner = "Algorithmic diagnosis"
) |>
  add_kappa_note("Quadratic weighted kappa", kappa_core_algo_algorithmic$quadratic_weighted_kappa) |>
  style_slide_gt()
```

## Three-class: HCAP Algo in Core vs Langa-Weir

```{r}
format_weighted_tab(
  tab_core_algo_lw,
  row_name = "core_algo",
  stubhead = "HCAP Algo in Core",
  spanner = "Langa-Weir diagnosis"
) |>
  add_kappa_note("Quadratic weighted kappa", kappa_core_algo_lw$quadratic_weighted_kappa) |>
  style_slide_gt()
```

## Three-class: Consensus vs PMM

```{r}
format_weighted_tab(
  tab_consensus_pmm,
  row_name = "consensus",
  stubhead = "Consensus panel diagnosis",
  spanner = "Profile-mixture classification"
) |>
  add_kappa_note("Quadratic weighted kappa", kappa_consensus_pmm$quadratic_weighted_kappa) |>
  style_slide_gt()
```

## Three-class: Algorithmic vs PMM

```{r}
format_weighted_tab(
  tab_algorithmic_pmm,
  row_name = "algorithmic",
  stubhead = "Algorithmic diagnosis",
  spanner = "Profile-mixture classification"
) |>
  add_kappa_note("Quadratic weighted kappa", kappa_algorithmic_pmm$quadratic_weighted_kappa) |>
  style_slide_gt()
```

## Three-class: Langa-Weir vs PMM

```{r}
format_weighted_tab(
  tab_lw_pmm,
  row_name = "langa_weir",
  stubhead = "Langa-Weir classification",
  spanner = "Profile-mixture classification"
) |>
  add_kappa_note("Quadratic weighted kappa", kappa_lw_pmm$quadratic_weighted_kappa) |>
  style_slide_gt()
```

## Three-class: HCAP Algo in Core vs PMM

```{r}
format_weighted_tab(
  tab_core_algo_pmm,
  row_name = "core_algo",
  stubhead = "HCAP Algo in Core",
  spanner = "Profile-mixture classification"
) |>
  add_kappa_note("Quadratic weighted kappa", kappa_core_algo_pmm$quadratic_weighted_kappa) |>
  style_slide_gt()
```

## Binary: Langa-Weir vs Consensus

```{r}
format_binary_weighted_tab(
  tab_lw_consensus_binary,
  row_name = "langa_weir_binary",
  stubhead = "Langa-Weir classification",
  spanner = "Consensus panel diagnosis"
) |>
  add_kappa_note("Kappa", kappa_lw_consensus_binary$kappa) |>
  style_slide_gt()
```

## Binary: Algorithmic vs Consensus

```{r}
format_binary_weighted_tab(
  tab_algorithmic_consensus_binary,
  row_name = "algorithmic_binary",
  stubhead = "Algorithmic diagnosis",
  spanner = "Consensus panel diagnosis"
) |>
  add_kappa_note("Kappa", kappa_algorithmic_consensus_binary$kappa) |>
  style_slide_gt()
```

## Binary: Algorithmic vs Langa-Weir

```{r}
format_binary_weighted_tab(
  tab_algorithmic_lw_binary,
  row_name = "algorithmic_binary",
  stubhead = "Algorithmic diagnosis",
  spanner = "Langa-Weir diagnosis"
) |>
  add_kappa_note("Kappa", kappa_algorithmic_lw_binary$kappa) |>
  style_slide_gt()
```

## Binary: HCAP Algo in Core vs Consensus

```{r}
format_binary_weighted_tab(
  tab_core_algo_consensus_binary,
  row_name = "core_algo_binary",
  stubhead = "HCAP Algo in Core",
  spanner = "Consensus panel diagnosis"
) |>
  add_kappa_note("Kappa", kappa_core_algo_consensus_binary$kappa) |>
  style_slide_gt()
```

## Binary: HCAP Algo in Core vs Algorithmic

```{r}
format_binary_weighted_tab(
  tab_core_algo_algorithmic_binary,
  row_name = "core_algo_binary",
  stubhead = "HCAP Algo in Core",
  spanner = "Algorithmic diagnosis"
) |>
  add_kappa_note("Kappa", kappa_core_algo_algorithmic_binary$kappa) |>
  style_slide_gt()
```

## Binary: HCAP Algo in Core vs Langa-Weir

```{r}
format_binary_weighted_tab(
  tab_core_algo_lw_binary,
  row_name = "core_algo_binary",
  stubhead = "HCAP Algo in Core",
  spanner = "Langa-Weir diagnosis"
) |>
  add_kappa_note("Kappa", kappa_core_algo_lw_binary$kappa) |>
  style_slide_gt()
```

## Binary: Consensus vs PMM

```{r}
format_binary_weighted_tab(
  tab_consensus_pmm_binary,
  row_name = "consensus_binary",
  stubhead = "Consensus panel diagnosis",
  spanner = "Profile-mixture classification"
) |>
  add_kappa_note("Kappa", kappa_consensus_pmm_binary$kappa) |>
  style_slide_gt()
```

## Binary: Algorithmic vs PMM

```{r}
format_binary_weighted_tab(
  tab_algorithmic_pmm_binary,
  row_name = "algorithmic_binary",
  stubhead = "Algorithmic diagnosis",
  spanner = "Profile-mixture classification"
) |>
  add_kappa_note("Kappa", kappa_algorithmic_pmm_binary$kappa) |>
  style_slide_gt()
```

## Binary: Langa-Weir vs PMM

```{r}
format_binary_weighted_tab(
  tab_lw_pmm_binary,
  row_name = "langa_weir_binary",
  stubhead = "Langa-Weir classification",
  spanner = "Profile-mixture classification"
) |>
  add_kappa_note("Kappa", kappa_lw_pmm_binary$kappa) |>
  style_slide_gt()
```

## Binary: HCAP Algo in Core vs PMM

```{r}
format_binary_weighted_tab(
  tab_core_algo_pmm_binary,
  row_name = "core_algo_binary",
  stubhead = "HCAP Algo in Core",
  spanner = "Profile-mixture classification"
) |>
  add_kappa_note("Kappa", kappa_core_algo_pmm_binary$kappa) |>
  style_slide_gt()
```